---
title: "Preprocessing for Clustering Analysis"
author: "Victoria Costas"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
---

```{r global_options, include=T, echo = F}
knitr::opts_chunk$set(echo = T, warning=FALSE, message=FALSE)
```

```{=html}
<style>
body {
text-align: justify}
</style>
```

# Preprocessing for clustering

Since the databases are not ready for analysis, certain changes will be made in order to have a database ready for the analysis we want to perform.

First of all, the "geo/time" column will be changed to "geo_time" so that there are no problems with special characters, and as this column is the one that contains the information on the countries (which is the one we will use), all the databases that do not contain it will be eliminated. Once this is done we will change ":" to NAs since eurostat uses ":" to indicate missing values, Eurostat also uses other characters to mark circumstances such as time jumps in values etc. and these characters will also be removed. In addition, columns that are not needed for the analysis and years before 2000 (due to lack of information, more than 80% of NAs) will be removed. Finally, we will create one new database combining all the previous datasets and make some final changes like converting to numerical or reordering the columns. 

**About the databases:**

Data was collected on 3rd of March 2023, thus if there have been any changes or updates to the databases after this date, they will not be taken into account. [Database](https://ec.europa.eu/eurostat/web/sdi/database) can be online accessed in the Eurostat web page. 


## Libraries

```{r}
library(tidyverse)
library(janitor)
library(dplyr)
library(mice)
```

## Read the data

```{r}
sdg_01.sdg_01_10 <- read_delim("data/sdg_01/new/sdg_01_10.tsv", delim = ",")
sdg_01.sdg_01_10a <- read_delim("data/sdg_01/new/sdg_01_10a.tsv", delim = ",")
sdg_01.sdg_01_20 <- read_delim("data/sdg_01/new/sdg_01_20.tsv", delim = ",")
sdg_01.sdg_01_20a <- read_delim("data/sdg_01/new/sdg_01_20a.tsv", delim = ",")
sdg_01.sdg_01_31 <- read_delim("data/sdg_01/new/sdg_01_31.tsv", delim = ",")
sdg_01.sdg_01_40 <- read_delim("data/sdg_01/new/sdg_01_40.tsv", delim = ",")
sdg_01.sdg_01_41 <- read_delim("data/sdg_01/new/sdg_01_41.tsv", delim = ",")
sdg_01.sdg_01_50 <- read_delim("data/sdg_01/new/sdg_01_50.tsv", delim = ",")
sdg_01.sdg_03_60 <- read_delim("data/sdg_01/new/sdg_03_60.tsv", delim = ",")
sdg_01.sdg_11_11 <- read_delim("data/sdg_01/new/sdg_11_11.tsv", delim = ",")
sdg_02.sdg_02_10 <- read_delim("data/sdg_02/new/sdg_02_10.tsv", delim = ",")
sdg_02.sdg_02_20 <- read_delim("data/sdg_02/new/sdg_02_20.tsv", delim = ",")
sdg_02.sdg_02_30 <- read_delim("data/sdg_02/new/sdg_02_30.tsv", delim = ",")
sdg_02.sdg_02_40 <- read_delim("data/sdg_02/new/sdg_02_40.tsv", delim = ",")
sdg_02.sdg_02_52 <- read_delim("data/sdg_02/new/sdg_02_52.tsv", delim = ",")
sdg_02.sdg_02_60 <- read_delim("data/sdg_02/new/sdg_02_60.tsv", delim = ",")
sdg_02.sdg_06_40 <- read_delim("data/sdg_02/new/sdg_06_40.tsv", delim = ",")
sdg_02.sdg_15_50 <- read_delim("data/sdg_02/new/sdg_15_50.tsv", delim = ",")
sdg_02.sdg_15_60 <- read_delim("data/sdg_02/new/sdg_15_60.tsv", delim = ",")
sdg_03.sdg_02_10 <- read_delim("data/sdg_03/new/sdg_02_10.tsv", delim = ",")
sdg_03.sdg_03_11 <- read_delim("data/sdg_03/new/sdg_03_11.tsv", delim = ",")
sdg_03.sdg_03_20 <- read_delim("data/sdg_03/new/sdg_03_20.tsv", delim = ",")
sdg_03.sdg_03_30 <- read_delim("data/sdg_03/new/sdg_03_30.tsv", delim = ",")
sdg_03.sdg_03_41 <- read_delim("data/sdg_03/new/sdg_03_41.tsv", delim = ",")
sdg_03.sdg_03_42 <- read_delim("data/sdg_03/new/sdg_03_42.tsv", delim = ",")
sdg_03.sdg_03_60 <- read_delim("data/sdg_03/new/sdg_03_60.tsv", delim = ",")
sdg_03.sdg_08_60 <- read_delim("data/sdg_03/new/sdg_08_60.tsv", delim = ",")
sdg_03.sdg_11_20 <- read_delim("data/sdg_03/new/sdg_11_20.tsv", delim = ",")
sdg_03.sdg_11_40 <- read_delim("data/sdg_03/new/sdg_11_40.tsv", delim = ",")
sdg_03.sdg_11_51 <- read_delim("data/sdg_03/new/sdg_11_51.tsv", delim = ",")
sdg_04.sdg_04_10 <- read_delim("data/sdg_04/new/sdg_04_10.tsv", delim = ",")
sdg_04.sdg_04_10a <- read_delim("data/sdg_04/new/sdg_04_10a.tsv", delim = ",")
sdg_04.sdg_04_20 <- read_delim("data/sdg_04/new/sdg_04_20.tsv", delim = ",")
sdg_04.sdg_04_31 <- read_delim("data/sdg_04/new/sdg_04_31.tsv", delim = ",")
sdg_04.sdg_04_40 <- read_delim("data/sdg_04/new/sdg_04_40.tsv", delim = ",")
sdg_04.sdg_04_60 <- read_delim("data/sdg_04/new/sdg_04_60.tsv", delim = ",")
sdg_04.sdg_04_70 <- read_delim("data/sdg_04/new/sdg_04_70.tsv", delim = ",")
sdg_05.sdg_04_10 <- read_delim("data/sdg_05/new/sdg_04_10.tsv", delim = ",")
sdg_05.sdg_04_20 <- read_delim("data/sdg_05/new/sdg_04_20.tsv", delim = ",")
sdg_05.sdg_05_10 <- read_delim("data/sdg_05/new/sdg_05_10.tsv", delim = ",")
sdg_05.sdg_05_20 <- read_delim("data/sdg_05/new/sdg_05_20.tsv", delim = ",")
sdg_05.sdg_05_30 <- read_delim("data/sdg_05/new/sdg_05_30.tsv", delim = ",")
sdg_05.sdg_05_40 <- read_delim("data/sdg_05/new/sdg_05_40.tsv", delim = ",")
sdg_05.sdg_05_50 <- read_delim("data/sdg_05/new/sdg_05_50.tsv", delim = ",")
sdg_05.sdg_05_60 <- read_delim("data/sdg_05/new/sdg_05_60.tsv", delim = ",")
sdg_06.sdg_06_10 <- read_delim("data/sdg_06/new/sdg_06_10.tsv", delim = ",")
sdg_06.sdg_06_20 <- read_delim("data/sdg_06/new/sdg_06_20.tsv", delim = ",")
sdg_06.sdg_06_30 <- read_delim("data/sdg_06/new/sdg_06_30.tsv", delim = ",")
sdg_06.sdg_06_40 <- read_delim("data/sdg_06/new/sdg_06_40.tsv", delim = ",")
sdg_06.sdg_06_50 <- read_delim("data/sdg_06/new/sdg_06_50.tsv", delim = ",")
sdg_06.sdg_06_60 <- read_delim("data/sdg_06/new/sdg_06_60.tsv", delim = ",")
sdg_06.sdg_14_40 <- read_delim("data/sdg_06/new/sdg_14_40.tsv", delim = ",")
sdg_07.sdg_07_10 <- read_delim("data/sdg_07/new/sdg_07_10.tsv", delim = ",")
sdg_07.sdg_07_11 <- read_delim("data/sdg_07/new/sdg_07_11.tsv", delim = ",")
sdg_07.sdg_07_20 <- read_delim("data/sdg_07/new/sdg_07_20.tsv", delim = ",")
sdg_07.sdg_07_30 <- read_delim("data/sdg_07/new/sdg_07_30.tsv", delim = ",")
sdg_07.sdg_07_40 <- read_delim("data/sdg_07/new/sdg_07_40.tsv", delim = ",")
sdg_07.sdg_07_50 <- read_delim("data/sdg_07/new/sdg_07_50.tsv", delim = ",")
sdg_07.sdg_07_60 <- read_delim("data/sdg_07/new/sdg_07_60.tsv", delim = ",")
sdg_08.sdg_01_41 <- read_delim("data/sdg_08/new/sdg_01_41.tsv", delim = ",")
sdg_08.sdg_05_40 <- read_delim("data/sdg_08/new/sdg_05_40.tsv", delim = ",")
sdg_08.sdg_08_10 <- read_delim("data/sdg_08/new/sdg_08_10.tsv", delim = ",")
sdg_08.sdg_08_11 <- read_delim("data/sdg_08/new/sdg_08_11.tsv", delim = ",")
sdg_08.sdg_08_20 <- read_delim("data/sdg_08/new/sdg_08_20.tsv", delim = ",")
sdg_08.sdg_08_20a <- read_delim("data/sdg_08/new/sdg_08_20a.tsv", delim = ",")
sdg_08.sdg_08_30 <- read_delim("data/sdg_08/new/sdg_08_30.tsv", delim = ",")
sdg_08.sdg_08_30a <- read_delim("data/sdg_08/new/sdg_08_30a.tsv", delim = ",")
sdg_08.sdg_08_40 <- read_delim("data/sdg_08/new/sdg_08_40.tsv", delim = ",")
sdg_08.sdg_08_60 <- read_delim("data/sdg_08/new/sdg_08_60.tsv", delim = ",")
sdg_09.sdg_04_20 <- read_delim("data/sdg_09/new/sdg_04_20.tsv", delim = ",")
sdg_09.sdg_09_10 <- read_delim("data/sdg_09/new/sdg_09_10.tsv", delim = ",")
sdg_09.sdg_09_30 <- read_delim("data/sdg_09/new/sdg_09_30.tsv", delim = ",")
sdg_09.sdg_09_40 <- read_delim("data/sdg_09/new/sdg_09_40.tsv", delim = ",")
sdg_09.sdg_09_50 <- read_delim("data/sdg_09/new/sdg_09_50.tsv", delim = ",")
sdg_09.sdg_09_60 <- read_delim("data/sdg_09/new/sdg_09_60.tsv", delim = ",")
sdg_09.sdg_09_70 <- read_delim("data/sdg_09/new/sdg_09_70.tsv", delim = ",")
sdg_09.sdg_12_61 <- read_delim("data/sdg_09/new/sdg_12_61.tsv", delim = ",")
sdg_09.sdg_17_60 <- read_delim("data/sdg_09/new/sdg_17_60.tsv", delim = ",")
sdg_10.sdg_01_10a <- read_delim("data/sdg_10/new/sdg_01_10a.tsv", delim = ",")
sdg_10.sdg_01_20a <- read_delim("data/sdg_10/new/sdg_01_20a.tsv", delim = ",")
sdg_10.sdg_04_10a <- read_delim("data/sdg_10/new/sdg_04_10a.tsv", delim = ",")
sdg_10.sdg_08_20a <- read_delim("data/sdg_10/new/sdg_08_20a.tsv", delim = ",")
sdg_10.sdg_08_30a <- read_delim("data/sdg_10/new/sdg_08_30a.tsv", delim = ",")
sdg_10.sdg_10_10 <- read_delim("data/sdg_10/new/sdg_10_10.tsv", delim = ",")
sdg_10.sdg_10_20 <- read_delim("data/sdg_10/new/sdg_10_20.tsv", delim = ",")
sdg_10.sdg_10_30 <- read_delim("data/sdg_10/new/sdg_10_30.tsv", delim = ",")
sdg_10.sdg_10_41 <- read_delim("data/sdg_10/new/sdg_10_41.tsv", delim = ",")
sdg_10.sdg_10_50 <- read_delim("data/sdg_10/new/sdg_10_50.tsv", delim = ",")
sdg_10.sdg_10_60 <- read_delim("data/sdg_10/new/sdg_10_60.tsv", delim = ",")
sdg_11.sdg_06_20 <- read_delim("data/sdg_11/new/sdg_06_20.tsv", delim = ",")
sdg_11.sdg_09_50 <- read_delim("data/sdg_11/new/sdg_09_50.tsv", delim = ",")
sdg_11.sdg_11_11 <- read_delim("data/sdg_11/new/sdg_11_11.tsv", delim = ",")
sdg_11.sdg_11_20 <- read_delim("data/sdg_11/new/sdg_11_20.tsv", delim = ",")
sdg_11.sdg_11_31 <- read_delim("data/sdg_11/new/sdg_11_31.tsv", delim = ",")
sdg_11.sdg_11_40 <- read_delim("data/sdg_11/new/sdg_11_40.tsv", delim = ",")
sdg_11.sdg_11_51 <- read_delim("data/sdg_11/new/sdg_11_51.tsv", delim = ",")
sdg_11.sdg_11_60 <- read_delim("data/sdg_11/new/sdg_11_60.tsv", delim = ",")
sdg_11.sdg_16_20 <- read_delim("data/sdg_11/new/sdg_16_20.tsv", delim = ",")
sdg_12.sdg_07_30 <- read_delim("data/sdg_12/new/sdg_07_30.tsv", delim = ",")
sdg_12.sdg_12_10 <- read_delim("data/sdg_12/new/sdg_12_10.tsv", delim = ",")
sdg_12.sdg_12_21 <- read_delim("data/sdg_12/new/sdg_12_21.tsv", delim = ",")
sdg_12.sdg_12_30 <- read_delim("data/sdg_12/new/sdg_12_30.tsv", delim = ",")
sdg_12.sdg_12_41 <- read_delim("data/sdg_12/new/sdg_12_41.tsv", delim = ",")
sdg_12.sdg_12_50 <- read_delim("data/sdg_12/new/sdg_12_50.tsv", delim = ",")
sdg_12.sdg_12_61 <- read_delim("data/sdg_12/new/sdg_12_61.tsv", delim = ",")
sdg_13.sdg_07_40 <- read_delim("data/sdg_13/new/sdg_07_40.tsv", delim = ",")
sdg_13.sdg_12_30 <- read_delim("data/sdg_13/new/sdg_12_30.tsv", delim = ",")
sdg_13.sdg_13_10 <- read_delim("data/sdg_13/new/sdg_13_10.tsv", delim = ",")
sdg_13.sdg_13_21 <- read_delim("data/sdg_13/new/sdg_13_21.tsv", delim = ",")
sdg_13.sdg_13_40 <- read_delim("data/sdg_13/new/sdg_13_40.tsv", delim = ",")
sdg_13.sdg_13_50 <- read_delim("data/sdg_13/new/sdg_13_50.tsv", delim = ",")
sdg_13.sdg_13_60 <- read_delim("data/sdg_13/new/sdg_13_60.tsv", delim = ",")
sdg_14.sdg_14_10 <- read_delim("data/sdg_14/new/sdg_14_10.tsv", delim = ",")
sdg_14.sdg_14_21 <- read_delim("data/sdg_14/new/sdg_14_21.tsv", delim = ",")
sdg_14.sdg_14_30 <- read_delim("data/sdg_14/new/sdg_14_30.tsv", delim = ",")
sdg_14.sdg_14_40 <- read_delim("data/sdg_14/new/sdg_14_40.tsv", delim = ",")
sdg_14.sdg_14_50 <- read_delim("data/sdg_14/new/sdg_14_50.tsv", delim = ",")
sdg_14.sdg_14_60 <- read_delim("data/sdg_14/new/sdg_14_60.tsv", delim = ",")
sdg_15.sdg_06_30 <- read_delim("data/sdg_15/new/sdg_06_30.tsv", delim = ",")
sdg_15.sdg_06_50 <- read_delim("data/sdg_15/new/sdg_06_50.tsv", delim = ",")
sdg_15.sdg_15_10 <- read_delim("data/sdg_15/new/sdg_15_10.tsv", delim = ",")
sdg_15.sdg_15_20 <- read_delim("data/sdg_15/new/sdg_15_20.tsv", delim = ",")
sdg_15.sdg_15_41 <- read_delim("data/sdg_15/new/sdg_15_41.tsv", delim = ",")
sdg_15.sdg_15_50 <- read_delim("data/sdg_15/new/sdg_15_50.tsv", delim = ",")
sdg_15.sdg_15_60 <- read_delim("data/sdg_15/new/sdg_15_60.tsv", delim = ",")
sdg_15.sdg_15_61 <- read_delim("data/sdg_15/new/sdg_15_61.tsv", delim = ",")
sdg_16.sdg_16_10 <- read_delim("data/sdg_16/new/sdg_16_10.tsv", delim = ",")
sdg_16.sdg_16_20 <- read_delim("data/sdg_16/new/sdg_16_20.tsv", delim = ",")
sdg_16.sdg_16_30 <- read_delim("data/sdg_16/new/sdg_16_30.tsv", delim = ",")
sdg_16.sdg_16_40 <- read_delim("data/sdg_16/new/sdg_16_40.tsv", delim = ",")
sdg_16.sdg_16_50 <- read_delim("data/sdg_16/new/sdg_16_50.tsv", delim = ",")
sdg_16.sdg_16_60 <- read_delim("data/sdg_16/new/sdg_16_60.tsv", delim = ",")
sdg_17.sdg_17_10 <- read_delim("data/sdg_17/new/sdg_17_10.tsv", delim = ",")
sdg_17.sdg_17_20 <- read_delim("data/sdg_17/new/sdg_17_20.tsv", delim = ",")
sdg_17.sdg_17_30 <- read_delim("data/sdg_17/new/sdg_17_30.tsv", delim = ",")
sdg_17.sdg_17_40 <- read_delim("data/sdg_17/new/sdg_17_40.tsv", delim = ",")
sdg_17.sdg_17_50 <- read_delim("data/sdg_17/new/sdg_17_50.tsv", delim = ",")
sdg_17.sdg_17_60 <- read_delim("data/sdg_17/new/sdg_17_60.tsv", delim = ",")
```

## Data cleaning and preprocessing

```{r}
#get a list of all database objects in the environment
database_names <- ls(pattern = "sdg_(.*)")

# --------------Change name from geo/time to geo_time
#function to change column name
change_column_name <- function(db_name) {
  db <- get(db_name)
  if ("geo\\time" %in% colnames(db)) {
    db <- db %>% rename(geo_time = `geo\\time`)
    assign(db_name, db, envir = .GlobalEnv)
    message(paste("Column name changed in", db_name))
  } else {
    message(paste("No column named 'geo\\time' found in", db_name))
  }
}

#apply column name change to all databases
lapply(database_names, change_column_name)
```

```{r}
#remove database without geo_time and select all the countries except from union of countries (EU, EA etc.)
remove_db_without_geo_time <- function(db_name) {
  db <- get(db_name)
  if ("geo_time" %in% colnames(db)) {
    db <- db %>% 
      filter(!grepl("^(EU|EA|COMM|EIB)", geo_time, ignore.case = TRUE))
    assign(db_name, db, envir = .GlobalEnv)
    message(paste("Filtered rows in", db_name))
  } else {
    message(paste("No column named 'geo_time' found in", db_name))
    rm(list = db_name, envir = .GlobalEnv)
    message(paste("Removed", db_name, "from the environment"))
  }
}

#apply column filtering to all the databases
lapply(database_names, remove_db_without_geo_time)
```


```{r}
#get a new list of all database objects in the environment
database_names <- ls(pattern = "sdg_(.*)")

#change ":" to missing values (NAs)
change_colon_to_na <- function(db_name) {
  db <- get(db_name)
  
  if (is.data.frame(db) || is.matrix(db)) {
    db[] <- lapply(db, function(x) ifelse(grepl(":", x), NA, x))
    assign(db_name, db, envir = .GlobalEnv)
    message(paste("Changed ':' to NA in", db_name))
  } else if (is.list(db)) {
    db <- lapply(db, function(x) {
      if (is.data.frame(x) || is.matrix(x)) {
        x[] <- lapply(x, function(y) ifelse(y == ":", NA, y))
        message(paste("Changed ':' to NA in a nested data frame or matrix within", db_name))
      }
      x
    })
    assign(db_name, db, envir = .GlobalEnv)
  } else {
    message(paste("Unsupported data structure found in", db_name))
  }
}

#apply column filtering to all the databases
lapply(database_names, change_colon_to_na)

```

There are some special values and flags that Eurostat uses to mark some data characteristics that need to be removed:

- b = break in time series
- c = confidential
- d = definition differs
- e = estimated
- f = forecast
- n = not significant
- p = provisional
- r = revised
- s = Eurostat estimate
- u = low reliability
- z = not applicable

```{r}
#remove flags
remove_chars_from_columns <- function(db_name) {
  db <- get(db_name)
  
  if (is.data.frame(db) || is.matrix(db)) {
    columns_to_process <- grep("^19|^20", colnames(db), value = TRUE)
    columns_to_process <- intersect(columns_to_process, colnames(db))
    
    db[, columns_to_process] <- lapply(db[, columns_to_process], function(x) gsub("[b-cdefnprsu-z]", "", x))
    
    assign(db_name, db, envir = .GlobalEnv)
    message(paste("Removed specified characters from columns starting with '19' or '20' in", db_name))
  } else if (is.list(db)) {
    db <- lapply(db, function(x) {
      if (is.data.frame(x) || is.matrix(x)) {
        columns_to_process <- grep("^19|^20", colnames(x), value = TRUE)
        columns_to_process <- intersect(columns_to_process, colnames(x))
        
        x[, columns_to_process] <- lapply(x[, columns_to_process], function(y) gsub("[b-cdefnprsu-z]", "", y))
        
        message(paste("Removed specified characters from columns starting with '19' or '20' in a nested data frame or matrix within", db_name))
      }
      x
    })
    assign(db_name, db, envir = .GlobalEnv)
  } else {
    message(paste("Unsupported data structure found in", db_name))
  }
}

#apply the function to all databases
lapply(database_names, remove_chars_from_columns)

```


```{r}
#remove non essential columns
remove_columns <- function(database_names) {
  for (database_name in database_names) {
    database <- get(database_name)

    # Identify columns to remove
    columns_to_remove <- c()
    for (col_name in colnames(database)) {
      if (!(col_name == "geo_time" || col_name == "indicator" || grepl("^20", col_name))) {
        columns_to_remove <- c(columns_to_remove, col_name)
      }
    }

    # Remove columns
    database <- database[, !(colnames(database) %in% columns_to_remove)]

    # Add updated database to global environment
    assign(database_name, database, envir = .GlobalEnv)
  }
}

#apply the function to all the databases
remove_columns(database_names)
```



```{r}
#convert to numeric
convert_to_numeric <- function(db) {
  db[, -1:-2] <- lapply(db[, -1:-2], as.numeric)
  return(db)
}

for (db_name in database_names) {
  if (exists(db_name)) {
    db <- get(db_name)
    if ("geo_time" %in% colnames(db)) {
      db <- convert_to_numeric(db)
      assign(db_name, db, envir = .GlobalEnv)
    } else {
      message(paste("No column named 'geo_time' found in", db_name))
    }
  }
}
```


```{r}
#keep only the years we are interested in
remove_years <- function(db_name) {
  if (exists(db_name)) {
    db <- get(db_name)
    db <- db %>%
      filter(year >= 2010 & year != 2022)
    assign(db_name, db, envir = .GlobalEnv)
    message(paste("Filtered years in", db_name))
  } else {
    message(paste("Database", db_name, "does not exist"))
  }
}

#initialize an empty data frame to store the merged data
merged_data <- data.frame()

#iterate over the database names
for (db_name in database_names) {
  if (exists(db_name)) {
    db <- get(db_name)
    merged_data <- bind_rows(merged_data, db)
  }
}

```


```{r}
#save the dataset
write.csv(merged_data, file = "./data/merged_data.csv")
```

