---
title: "Preprocessing for Simple Trend Analysis"
author: "Victoria Costas"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
---

```{r global_options, include=T, echo = F}
knitr::opts_chunk$set(echo = T, warning=FALSE, message=FALSE)
```

```{=html}
<style>
body {
text-align: justify}
</style>
```

# Preprocessing for trends 

To perform the preprocessing of the data for the simple trend analysis, we will use the same strategy as for the preprocessing of the data for the clustering but this time we will only keep the rows for Spain and Germany and we will save the dataset in long format in order to facilitate analysis. 

## Libraries

```{r}
library(tidyverse)
library(janitor)
library(dplyr)
library(mice)
```

## Read the data

```{r}
sdg_01.sdg_01_10 <- read_delim("data/sdg_01/new/sdg_01_10.tsv", delim = ",")
sdg_01.sdg_01_10a <- read_delim("data/sdg_01/new/sdg_01_10a.tsv", delim = ",")
sdg_01.sdg_01_20 <- read_delim("data/sdg_01/new/sdg_01_20.tsv", delim = ",")
sdg_01.sdg_01_20a <- read_delim("data/sdg_01/new/sdg_01_20a.tsv", delim = ",")
sdg_01.sdg_01_31 <- read_delim("data/sdg_01/new/sdg_01_31.tsv", delim = ",")
sdg_01.sdg_01_40 <- read_delim("data/sdg_01/new/sdg_01_40.tsv", delim = ",")
sdg_01.sdg_01_41 <- read_delim("data/sdg_01/new/sdg_01_41.tsv", delim = ",")
sdg_01.sdg_01_50 <- read_delim("data/sdg_01/new/sdg_01_50.tsv", delim = ",")
sdg_01.sdg_03_60 <- read_delim("data/sdg_01/new/sdg_03_60.tsv", delim = ",")
sdg_01.sdg_11_11 <- read_delim("data/sdg_01/new/sdg_11_11.tsv", delim = ",")
sdg_02.sdg_02_10 <- read_delim("data/sdg_02/new/sdg_02_10.tsv", delim = ",")
sdg_02.sdg_02_20 <- read_delim("data/sdg_02/new/sdg_02_20.tsv", delim = ",")
sdg_02.sdg_02_30 <- read_delim("data/sdg_02/new/sdg_02_30.tsv", delim = ",")
sdg_02.sdg_02_40 <- read_delim("data/sdg_02/new/sdg_02_40.tsv", delim = ",")
sdg_02.sdg_02_52 <- read_delim("data/sdg_02/new/sdg_02_52.tsv", delim = ",")
sdg_02.sdg_02_60 <- read_delim("data/sdg_02/new/sdg_02_60.tsv", delim = ",")
sdg_02.sdg_06_40 <- read_delim("data/sdg_02/new/sdg_06_40.tsv", delim = ",")
sdg_02.sdg_15_50 <- read_delim("data/sdg_02/new/sdg_15_50.tsv", delim = ",")
sdg_02.sdg_15_60 <- read_delim("data/sdg_02/new/sdg_15_60.tsv", delim = ",")
sdg_03.sdg_02_10 <- read_delim("data/sdg_03/new/sdg_02_10.tsv", delim = ",")
sdg_03.sdg_03_11 <- read_delim("data/sdg_03/new/sdg_03_11.tsv", delim = ",")
sdg_03.sdg_03_20 <- read_delim("data/sdg_03/new/sdg_03_20.tsv", delim = ",")
sdg_03.sdg_03_30 <- read_delim("data/sdg_03/new/sdg_03_30.tsv", delim = ",")
sdg_03.sdg_03_41 <- read_delim("data/sdg_03/new/sdg_03_41.tsv", delim = ",")
sdg_03.sdg_03_42 <- read_delim("data/sdg_03/new/sdg_03_42.tsv", delim = ",")
sdg_03.sdg_03_60 <- read_delim("data/sdg_03/new/sdg_03_60.tsv", delim = ",")
sdg_03.sdg_08_60 <- read_delim("data/sdg_03/new/sdg_08_60.tsv", delim = ",")
sdg_03.sdg_11_20 <- read_delim("data/sdg_03/new/sdg_11_20.tsv", delim = ",")
sdg_03.sdg_11_40 <- read_delim("data/sdg_03/new/sdg_11_40.tsv", delim = ",")
sdg_03.sdg_11_51 <- read_delim("data/sdg_03/new/sdg_11_51.tsv", delim = ",")
sdg_04.sdg_04_10 <- read_delim("data/sdg_04/new/sdg_04_10.tsv", delim = ",")
sdg_04.sdg_04_10a <- read_delim("data/sdg_04/new/sdg_04_10a.tsv", delim = ",")
sdg_04.sdg_04_20 <- read_delim("data/sdg_04/new/sdg_04_20.tsv", delim = ",")
sdg_04.sdg_04_31 <- read_delim("data/sdg_04/new/sdg_04_31.tsv", delim = ",")
sdg_04.sdg_04_40 <- read_delim("data/sdg_04/new/sdg_04_40.tsv", delim = ",")
sdg_04.sdg_04_60 <- read_delim("data/sdg_04/new/sdg_04_60.tsv", delim = ",")
sdg_04.sdg_04_70 <- read_delim("data/sdg_04/new/sdg_04_70.tsv", delim = ",")
sdg_05.sdg_04_10 <- read_delim("data/sdg_05/new/sdg_04_10.tsv", delim = ",")
sdg_05.sdg_04_20 <- read_delim("data/sdg_05/new/sdg_04_20.tsv", delim = ",")
sdg_05.sdg_05_10 <- read_delim("data/sdg_05/new/sdg_05_10.tsv", delim = ",")
sdg_05.sdg_05_20 <- read_delim("data/sdg_05/new/sdg_05_20.tsv", delim = ",")
sdg_05.sdg_05_30 <- read_delim("data/sdg_05/new/sdg_05_30.tsv", delim = ",")
sdg_05.sdg_05_40 <- read_delim("data/sdg_05/new/sdg_05_40.tsv", delim = ",")
sdg_05.sdg_05_50 <- read_delim("data/sdg_05/new/sdg_05_50.tsv", delim = ",")
sdg_05.sdg_05_60 <- read_delim("data/sdg_05/new/sdg_05_60.tsv", delim = ",")
sdg_06.sdg_06_10 <- read_delim("data/sdg_06/new/sdg_06_10.tsv", delim = ",")
sdg_06.sdg_06_20 <- read_delim("data/sdg_06/new/sdg_06_20.tsv", delim = ",")
sdg_06.sdg_06_30 <- read_delim("data/sdg_06/new/sdg_06_30.tsv", delim = ",")
sdg_06.sdg_06_40 <- read_delim("data/sdg_06/new/sdg_06_40.tsv", delim = ",")
sdg_06.sdg_06_50 <- read_delim("data/sdg_06/new/sdg_06_50.tsv", delim = ",")
sdg_06.sdg_06_60 <- read_delim("data/sdg_06/new/sdg_06_60.tsv", delim = ",")
sdg_06.sdg_14_40 <- read_delim("data/sdg_06/new/sdg_14_40.tsv", delim = ",")
sdg_07.sdg_07_10 <- read_delim("data/sdg_07/new/sdg_07_10.tsv", delim = ",")
sdg_07.sdg_07_11 <- read_delim("data/sdg_07/new/sdg_07_11.tsv", delim = ",")
sdg_07.sdg_07_20 <- read_delim("data/sdg_07/new/sdg_07_20.tsv", delim = ",")
sdg_07.sdg_07_30 <- read_delim("data/sdg_07/new/sdg_07_30.tsv", delim = ",")
sdg_07.sdg_07_40 <- read_delim("data/sdg_07/new/sdg_07_40.tsv", delim = ",")
sdg_07.sdg_07_50 <- read_delim("data/sdg_07/new/sdg_07_50.tsv", delim = ",")
sdg_07.sdg_07_60 <- read_delim("data/sdg_07/new/sdg_07_60.tsv", delim = ",")
sdg_08.sdg_01_41 <- read_delim("data/sdg_08/new/sdg_01_41.tsv", delim = ",")
sdg_08.sdg_05_40 <- read_delim("data/sdg_08/new/sdg_05_40.tsv", delim = ",")
sdg_08.sdg_08_10 <- read_delim("data/sdg_08/new/sdg_08_10.tsv", delim = ",")
sdg_08.sdg_08_11 <- read_delim("data/sdg_08/new/sdg_08_11.tsv", delim = ",")
sdg_08.sdg_08_20 <- read_delim("data/sdg_08/new/sdg_08_20.tsv", delim = ",")
sdg_08.sdg_08_20a <- read_delim("data/sdg_08/new/sdg_08_20a.tsv", delim = ",")
sdg_08.sdg_08_30 <- read_delim("data/sdg_08/new/sdg_08_30.tsv", delim = ",")
sdg_08.sdg_08_30a <- read_delim("data/sdg_08/new/sdg_08_30a.tsv", delim = ",")
sdg_08.sdg_08_40 <- read_delim("data/sdg_08/new/sdg_08_40.tsv", delim = ",")
sdg_08.sdg_08_60 <- read_delim("data/sdg_08/new/sdg_08_60.tsv", delim = ",")
sdg_09.sdg_04_20 <- read_delim("data/sdg_09/new/sdg_04_20.tsv", delim = ",")
sdg_09.sdg_09_10 <- read_delim("data/sdg_09/new/sdg_09_10.tsv", delim = ",")
sdg_09.sdg_09_30 <- read_delim("data/sdg_09/new/sdg_09_30.tsv", delim = ",")
sdg_09.sdg_09_40 <- read_delim("data/sdg_09/new/sdg_09_40.tsv", delim = ",")
sdg_09.sdg_09_50 <- read_delim("data/sdg_09/new/sdg_09_50.tsv", delim = ",")
sdg_09.sdg_09_60 <- read_delim("data/sdg_09/new/sdg_09_60.tsv", delim = ",")
sdg_09.sdg_09_70 <- read_delim("data/sdg_09/new/sdg_09_70.tsv", delim = ",")
sdg_09.sdg_12_61 <- read_delim("data/sdg_09/new/sdg_12_61.tsv", delim = ",")
sdg_09.sdg_17_60 <- read_delim("data/sdg_09/new/sdg_17_60.tsv", delim = ",")
sdg_10.sdg_01_10a <- read_delim("data/sdg_10/new/sdg_01_10a.tsv", delim = ",")
sdg_10.sdg_01_20a <- read_delim("data/sdg_10/new/sdg_01_20a.tsv", delim = ",")
sdg_10.sdg_04_10a <- read_delim("data/sdg_10/new/sdg_04_10a.tsv", delim = ",")
sdg_10.sdg_08_20a <- read_delim("data/sdg_10/new/sdg_08_20a.tsv", delim = ",")
sdg_10.sdg_08_30a <- read_delim("data/sdg_10/new/sdg_08_30a.tsv", delim = ",")
sdg_10.sdg_10_10 <- read_delim("data/sdg_10/new/sdg_10_10.tsv", delim = ",")
sdg_10.sdg_10_20 <- read_delim("data/sdg_10/new/sdg_10_20.tsv", delim = ",")
sdg_10.sdg_10_30 <- read_delim("data/sdg_10/new/sdg_10_30.tsv", delim = ",")
sdg_10.sdg_10_41 <- read_delim("data/sdg_10/new/sdg_10_41.tsv", delim = ",")
sdg_10.sdg_10_50 <- read_delim("data/sdg_10/new/sdg_10_50.tsv", delim = ",")
sdg_10.sdg_10_60 <- read_delim("data/sdg_10/new/sdg_10_60.tsv", delim = ",")
sdg_11.sdg_06_20 <- read_delim("data/sdg_11/new/sdg_06_20.tsv", delim = ",")
sdg_11.sdg_09_50 <- read_delim("data/sdg_11/new/sdg_09_50.tsv", delim = ",")
sdg_11.sdg_11_11 <- read_delim("data/sdg_11/new/sdg_11_11.tsv", delim = ",")
sdg_11.sdg_11_20 <- read_delim("data/sdg_11/new/sdg_11_20.tsv", delim = ",")
sdg_11.sdg_11_31 <- read_delim("data/sdg_11/new/sdg_11_31.tsv", delim = ",")
sdg_11.sdg_11_40 <- read_delim("data/sdg_11/new/sdg_11_40.tsv", delim = ",")
sdg_11.sdg_11_51 <- read_delim("data/sdg_11/new/sdg_11_51.tsv", delim = ",")
sdg_11.sdg_11_60 <- read_delim("data/sdg_11/new/sdg_11_60.tsv", delim = ",")
sdg_11.sdg_16_20 <- read_delim("data/sdg_11/new/sdg_16_20.tsv", delim = ",")
sdg_12.sdg_07_30 <- read_delim("data/sdg_12/new/sdg_07_30.tsv", delim = ",")
sdg_12.sdg_12_10 <- read_delim("data/sdg_12/new/sdg_12_10.tsv", delim = ",")
sdg_12.sdg_12_21 <- read_delim("data/sdg_12/new/sdg_12_21.tsv", delim = ",")
sdg_12.sdg_12_30 <- read_delim("data/sdg_12/new/sdg_12_30.tsv", delim = ",")
sdg_12.sdg_12_41 <- read_delim("data/sdg_12/new/sdg_12_41.tsv", delim = ",")
sdg_12.sdg_12_50 <- read_delim("data/sdg_12/new/sdg_12_50.tsv", delim = ",")
sdg_12.sdg_12_61 <- read_delim("data/sdg_12/new/sdg_12_61.tsv", delim = ",")
sdg_13.sdg_07_40 <- read_delim("data/sdg_13/new/sdg_07_40.tsv", delim = ",")
sdg_13.sdg_12_30 <- read_delim("data/sdg_13/new/sdg_12_30.tsv", delim = ",")
sdg_13.sdg_13_10 <- read_delim("data/sdg_13/new/sdg_13_10.tsv", delim = ",")
sdg_13.sdg_13_21 <- read_delim("data/sdg_13/new/sdg_13_21.tsv", delim = ",")
sdg_13.sdg_13_40 <- read_delim("data/sdg_13/new/sdg_13_40.tsv", delim = ",")
sdg_13.sdg_13_50 <- read_delim("data/sdg_13/new/sdg_13_50.tsv", delim = ",")
sdg_13.sdg_13_60 <- read_delim("data/sdg_13/new/sdg_13_60.tsv", delim = ",")
sdg_14.sdg_14_10 <- read_delim("data/sdg_14/new/sdg_14_10.tsv", delim = ",")
sdg_14.sdg_14_21 <- read_delim("data/sdg_14/new/sdg_14_21.tsv", delim = ",")
sdg_14.sdg_14_30 <- read_delim("data/sdg_14/new/sdg_14_30.tsv", delim = ",")
sdg_14.sdg_14_40 <- read_delim("data/sdg_14/new/sdg_14_40.tsv", delim = ",")
sdg_14.sdg_14_50 <- read_delim("data/sdg_14/new/sdg_14_50.tsv", delim = ",")
sdg_14.sdg_14_60 <- read_delim("data/sdg_14/new/sdg_14_60.tsv", delim = ",")
sdg_15.sdg_06_30 <- read_delim("data/sdg_15/new/sdg_06_30.tsv", delim = ",")
sdg_15.sdg_06_50 <- read_delim("data/sdg_15/new/sdg_06_50.tsv", delim = ",")
sdg_15.sdg_15_10 <- read_delim("data/sdg_15/new/sdg_15_10.tsv", delim = ",")
sdg_15.sdg_15_20 <- read_delim("data/sdg_15/new/sdg_15_20.tsv", delim = ",")
sdg_15.sdg_15_41 <- read_delim("data/sdg_15/new/sdg_15_41.tsv", delim = ",")
sdg_15.sdg_15_50 <- read_delim("data/sdg_15/new/sdg_15_50.tsv", delim = ",")
sdg_15.sdg_15_60 <- read_delim("data/sdg_15/new/sdg_15_60.tsv", delim = ",")
sdg_15.sdg_15_61 <- read_delim("data/sdg_15/new/sdg_15_61.tsv", delim = ",")
sdg_16.sdg_16_10 <- read_delim("data/sdg_16/new/sdg_16_10.tsv", delim = ",")
sdg_16.sdg_16_20 <- read_delim("data/sdg_16/new/sdg_16_20.tsv", delim = ",")
sdg_16.sdg_16_30 <- read_delim("data/sdg_16/new/sdg_16_30.tsv", delim = ",")
sdg_16.sdg_16_40 <- read_delim("data/sdg_16/new/sdg_16_40.tsv", delim = ",")
sdg_16.sdg_16_50 <- read_delim("data/sdg_16/new/sdg_16_50.tsv", delim = ",")
sdg_16.sdg_16_60 <- read_delim("data/sdg_16/new/sdg_16_60.tsv", delim = ",")
sdg_17.sdg_17_10 <- read_delim("data/sdg_17/new/sdg_17_10.tsv", delim = ",")
sdg_17.sdg_17_20 <- read_delim("data/sdg_17/new/sdg_17_20.tsv", delim = ",")
sdg_17.sdg_17_30 <- read_delim("data/sdg_17/new/sdg_17_30.tsv", delim = ",")
sdg_17.sdg_17_40 <- read_delim("data/sdg_17/new/sdg_17_40.tsv", delim = ",")
sdg_17.sdg_17_50 <- read_delim("data/sdg_17/new/sdg_17_50.tsv", delim = ",")
sdg_17.sdg_17_60 <- read_delim("data/sdg_17/new/sdg_17_60.tsv", delim = ",")


```

## Data cleaning and preprocessing

```{r}
#get a list of all database objects in the environment
database_names <- ls(pattern = "sdg_(.*)")

# --------------Change name from geo/time to geo_time
# Function to change column name
change_column_name <- function(db_name) {
  db <- get(db_name)
  if ("geo\\time" %in% colnames(db)) {
    db <- db %>% rename(geo_time = `geo\\time`)
    assign(db_name, db, envir = .GlobalEnv)
    message(paste("Column name changed in", db_name))
  } else {
    message(paste("No column named 'geo\\time' found in", db_name))
  }
}

#apply column name change to all databases
lapply(database_names, change_column_name)
```

```{r}
#remove database without geo_time and select Spain and Germany
remove_db_without_geo_time <- function(db_name) {
  db <- get(db_name)
  if ("geo_time" %in% colnames(db)) {
    db <- db %>% 
      filter(geo_time %in% c("ES", "DE"))
    assign(db_name, db, envir = .GlobalEnv)
    message(paste("Filtered rows in", db_name))
  } else {
    message(paste("No column named 'geo_time' found in", db_name))
    rm(list = db_name, envir = .GlobalEnv)
    message(paste("Removed", db_name, "from the environment"))
  }
}

#apply column filtering to all the databases
lapply(database_names, remove_db_without_geo_time)
```


```{r}
#get a new list of all database objects in the environment
database_names <- ls(pattern = "sdg_(.*)")

#change ":" to missing values (NAs)
change_colon_to_na <- function(db_name) {
  db <- get(db_name)
  
  if (is.data.frame(db) || is.matrix(db)) {
    db[] <- lapply(db, function(x) ifelse(grepl(":", x), NA, x))
    assign(db_name, db, envir = .GlobalEnv)
    message(paste("Changed ':' to NA in", db_name))
  } else if (is.list(db)) {
    db <- lapply(db, function(x) {
      if (is.data.frame(x) || is.matrix(x)) {
        x[] <- lapply(x, function(y) ifelse(y == ":", NA, y))
        message(paste("Changed ':' to NA in a nested data frame or matrix within", db_name))
      }
      x
    })
    assign(db_name, db, envir = .GlobalEnv)
  } else {
    message(paste("Unsupported data structure found in", db_name))
  }
}

#apply column filtering to all the databases
lapply(database_names, change_colon_to_na)

```


```{r}
#remove the flags
remove_chars_from_columns <- function(db_name) {
  db <- get(db_name)
  
  if (is.data.frame(db) || is.matrix(db)) {
    columns_to_process <- grep("^19|^20", colnames(db), value = TRUE)
    columns_to_process <- intersect(columns_to_process, colnames(db))
    
    db[, columns_to_process] <- lapply(db[, columns_to_process], function(x) gsub("[b-cdefnprsu-z]", "", x))
    
    assign(db_name, db, envir = .GlobalEnv)
    message(paste("Removed specified characters from columns starting with '19' or '20' in", db_name))
  } else if (is.list(db)) {
    db <- lapply(db, function(x) {
      if (is.data.frame(x) || is.matrix(x)) {
        columns_to_process <- grep("^19|^20", colnames(x), value = TRUE)
        columns_to_process <- intersect(columns_to_process, colnames(x))
        
        x[, columns_to_process] <- lapply(x[, columns_to_process], function(y) gsub("[b-cdefnprsu-z]", "", y))
        
        message(paste("Removed specified characters from columns starting with '19' or '20' in a nested data frame or matrix within", db_name))
      }
      x
    })
    assign(db_name, db, envir = .GlobalEnv)
  } else {
    message(paste("Unsupported data structure found in", db_name))
  }
}

#apply the function to all databases
lapply(database_names, remove_chars_from_columns)

```


```{r}
#remove non essential columns
remove_columns <- function(database_names) {
  for (database_name in database_names) {
    database <- get(database_name)

    #identify columns to remove
    columns_to_remove <- c()
    for (col_name in colnames(database)) {
      if (!(col_name == "geo_time" || col_name == "indicator" || grepl("^20", col_name))) {
        columns_to_remove <- c(columns_to_remove, col_name)
      }
    }

    #remove columns
    database <- database[, !(colnames(database) %in% columns_to_remove)]

    assign(database_name, database, envir = .GlobalEnv)
  }
}
#apply the function to all the databases
remove_columns(database_names)
```



```{r}
# Create dataset for Spain
spain_dataset <- data.frame()
spain_database_names <- ls(pattern = "^sdg_")
for (database_name in spain_database_names) {
  database <- get(database_name)
  database$geo_time <- as.character(database$geo_time)  # Convert geo_time column to character
  database$indicator <- as.character(database$indicator) 
  spain_rows <- database[database$geo_time == "ES", ]
  spain_dataset <- bind_rows(spain_dataset, spain_rows)
}

# Create dataset for Germany
germany_dataset <- data.frame()
germany_database_names <- ls(pattern = "^sdg_")
for (database_name in germany_database_names) {
  database <- get(database_name)
  database$geo_time <- as.character(database$geo_time)  # Convert geo_time column to character
  database$indicator <- as.character(database$indicator)
  germany_rows <- database[database$geo_time == "DE", ]
  germany_dataset <- bind_rows(germany_dataset, germany_rows)
}

```


```{r}
#convert to numeric
spain_dataset[, -1:-2] <- lapply(spain_dataset[, -1:-2], function(x) as.numeric(as.character(x)))
germany_dataset[, -1:-2] <- lapply(germany_dataset[, -1:-2], function(x) as.numeric(as.character(x)))
```


```{r}
#remove the column for 2022 in Spain dataset
spain_dataset <- spain_dataset %>% select(-`2022 `)
#remove the column for 2022 in Germany dataset
germany_dataset <- germany_dataset %>% select(-`2022 `)

```


```{r}
#reorder columns in Spain dataset in reverse chronological order
spain_dataset <- spain_dataset[, rev(order(colnames(spain_dataset)))]
#reorder columns in Germany dataset in reverse chronological order
germany_dataset <- germany_dataset[, rev(order(colnames(germany_dataset)))]
```


```{r}
#creating the new datasets
spain_trends <- spain_dataset %>%
  pivot_longer(cols = starts_with("20"), 
               names_to = "year", 
               values_to = "value")

germany_trends <- germany_dataset %>%
  pivot_longer(cols = starts_with("20"), 
               names_to = "year", 
               values_to = "value")
```


```{r}
#save the datasets 
write.csv(spain_trends, file = "./data/spain_trends.csv")
write.csv(germany_trends, file = "./data/germany_trends.csv")
```

