---
title: "Draft TFM"
author: "Victoria Costas"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
---

```{r global_options, include=T, echo = F}
knitr::opts_chunk$set(echo = T, warning=FALSE, message=FALSE)
```

```{=html}
<style>
body {
text-align: justify}
</style>
```

# Preprocessing for trends 

To perform the pre-processing of the data for the simple trend analysis, we will use the same strategy as for the pre-processing of the data for the clustering but this time in addition to the column indicating the country we will also keep the column indicating the target to which each indicator belongs and apply the long format which is more appropriate for this analysis.

**Note** that I have added a new column named indicator to point out which indicator belongs to which goal, due to reasons of time and simplicity this was done with bash.

## Libraries

```{r}
library(tidyverse)
library(janitor)
library(dplyr)
library(mice)
```

## Read the data

Data collected on the 3rd of March 2023.

```{r}
sdg_01.sdg_01_10 <- read_delim("data/sdg_01/new/sdg_01_10.tsv", delim = ",")
sdg_01.sdg_01_10a <- read_delim("data/sdg_01/new/sdg_01_10a.tsv", delim = ",")
sdg_01.sdg_01_20 <- read_delim("data/sdg_01/new/sdg_01_20.tsv", delim = ",")
sdg_01.sdg_01_20a <- read_delim("data/sdg_01/new/sdg_01_20a.tsv", delim = ",")
sdg_01.sdg_01_31 <- read_delim("data/sdg_01/new/sdg_01_31.tsv", delim = ",")
sdg_01.sdg_01_40 <- read_delim("data/sdg_01/new/sdg_01_40.tsv", delim = ",")
sdg_01.sdg_01_41 <- read_delim("data/sdg_01/new/sdg_01_41.tsv", delim = ",")
sdg_01.sdg_01_50 <- read_delim("data/sdg_01/new/sdg_01_50.tsv", delim = ",")
sdg_01.sdg_03_60 <- read_delim("data/sdg_01/new/sdg_03_60.tsv", delim = ",")
sdg_01.sdg_11_11 <- read_delim("data/sdg_01/new/sdg_11_11.tsv", delim = ",")
sdg_02.sdg_02_10 <- read_delim("data/sdg_02/new/sdg_02_10.tsv", delim = ",")
sdg_02.sdg_02_20 <- read_delim("data/sdg_02/new/sdg_02_20.tsv", delim = ",")
sdg_02.sdg_02_30 <- read_delim("data/sdg_02/new/sdg_02_30.tsv", delim = ",")
sdg_02.sdg_02_40 <- read_delim("data/sdg_02/new/sdg_02_40.tsv", delim = ",")
sdg_02.sdg_02_52 <- read_delim("data/sdg_02/new/sdg_02_52.tsv", delim = ",")
sdg_02.sdg_02_60 <- read_delim("data/sdg_02/new/sdg_02_60.tsv", delim = ",")
sdg_02.sdg_06_40 <- read_delim("data/sdg_02/new/sdg_06_40.tsv", delim = ",")
sdg_02.sdg_15_50 <- read_delim("data/sdg_02/new/sdg_15_50.tsv", delim = ",")
sdg_02.sdg_15_60 <- read_delim("data/sdg_02/new/sdg_15_60.tsv", delim = ",")
sdg_03.sdg_02_10 <- read_delim("data/sdg_03/new/sdg_02_10.tsv", delim = ",")
sdg_03.sdg_03_11 <- read_delim("data/sdg_03/new/sdg_03_11.tsv", delim = ",")
sdg_03.sdg_03_20 <- read_delim("data/sdg_03/new/sdg_03_20.tsv", delim = ",")
sdg_03.sdg_03_30 <- read_delim("data/sdg_03/new/sdg_03_30.tsv", delim = ",")
sdg_03.sdg_03_41 <- read_delim("data/sdg_03/new/sdg_03_41.tsv", delim = ",")
sdg_03.sdg_03_42 <- read_delim("data/sdg_03/new/sdg_03_42.tsv", delim = ",")
sdg_03.sdg_03_60 <- read_delim("data/sdg_03/new/sdg_03_60.tsv", delim = ",")
sdg_03.sdg_08_60 <- read_delim("data/sdg_03/new/sdg_08_60.tsv", delim = ",")
sdg_03.sdg_11_20 <- read_delim("data/sdg_03/new/sdg_11_20.tsv", delim = ",")
sdg_03.sdg_11_40 <- read_delim("data/sdg_03/new/sdg_11_40.tsv", delim = ",")
sdg_03.sdg_11_51 <- read_delim("data/sdg_03/new/sdg_11_51.tsv", delim = ",")
sdg_04.sdg_04_10 <- read_delim("data/sdg_04/new/sdg_04_10.tsv", delim = ",")
sdg_04.sdg_04_10a <- read_delim("data/sdg_04/new/sdg_04_10a.tsv", delim = ",")
sdg_04.sdg_04_20 <- read_delim("data/sdg_04/new/sdg_04_20.tsv", delim = ",")
sdg_04.sdg_04_31 <- read_delim("data/sdg_04/new/sdg_04_31.tsv", delim = ",")
sdg_04.sdg_04_40 <- read_delim("data/sdg_04/new/sdg_04_40.tsv", delim = ",")
sdg_04.sdg_04_60 <- read_delim("data/sdg_04/new/sdg_04_60.tsv", delim = ",")
sdg_04.sdg_04_70 <- read_delim("data/sdg_04/new/sdg_04_70.tsv", delim = ",")
sdg_05.sdg_04_10 <- read_delim("data/sdg_05/new/sdg_04_10.tsv", delim = ",")
sdg_05.sdg_04_20 <- read_delim("data/sdg_05/new/sdg_04_20.tsv", delim = ",")
sdg_05.sdg_05_10 <- read_delim("data/sdg_05/new/sdg_05_10.tsv", delim = ",")
sdg_05.sdg_05_20 <- read_delim("data/sdg_05/new/sdg_05_20.tsv", delim = ",")
sdg_05.sdg_05_30 <- read_delim("data/sdg_05/new/sdg_05_30.tsv", delim = ",")
sdg_05.sdg_05_40 <- read_delim("data/sdg_05/new/sdg_05_40.tsv", delim = ",")
sdg_05.sdg_05_50 <- read_delim("data/sdg_05/new/sdg_05_50.tsv", delim = ",")
sdg_05.sdg_05_60 <- read_delim("data/sdg_05/new/sdg_05_60.tsv", delim = ",")
sdg_06.sdg_06_10 <- read_delim("data/sdg_06/new/sdg_06_10.tsv", delim = ",")
sdg_06.sdg_06_20 <- read_delim("data/sdg_06/new/sdg_06_20.tsv", delim = ",")
sdg_06.sdg_06_30 <- read_delim("data/sdg_06/new/sdg_06_30.tsv", delim = ",")
sdg_06.sdg_06_40 <- read_delim("data/sdg_06/new/sdg_06_40.tsv", delim = ",")
sdg_06.sdg_06_50 <- read_delim("data/sdg_06/new/sdg_06_50.tsv", delim = ",")
sdg_06.sdg_06_60 <- read_delim("data/sdg_06/new/sdg_06_60.tsv", delim = ",")
sdg_06.sdg_14_40 <- read_delim("data/sdg_06/new/sdg_14_40.tsv", delim = ",")
sdg_07.sdg_07_10 <- read_delim("data/sdg_07/new/sdg_07_10.tsv", delim = ",")
sdg_07.sdg_07_11 <- read_delim("data/sdg_07/new/sdg_07_11.tsv", delim = ",")
sdg_07.sdg_07_20 <- read_delim("data/sdg_07/new/sdg_07_20.tsv", delim = ",")
sdg_07.sdg_07_30 <- read_delim("data/sdg_07/new/sdg_07_30.tsv", delim = ",")
sdg_07.sdg_07_40 <- read_delim("data/sdg_07/new/sdg_07_40.tsv", delim = ",")
sdg_07.sdg_07_50 <- read_delim("data/sdg_07/new/sdg_07_50.tsv", delim = ",")
sdg_07.sdg_07_60 <- read_delim("data/sdg_07/new/sdg_07_60.tsv", delim = ",")
sdg_08.sdg_01_41 <- read_delim("data/sdg_08/new/sdg_01_41.tsv", delim = ",")
sdg_08.sdg_05_40 <- read_delim("data/sdg_08/new/sdg_05_40.tsv", delim = ",")
sdg_08.sdg_08_10 <- read_delim("data/sdg_08/new/sdg_08_10.tsv", delim = ",")
sdg_08.sdg_08_11 <- read_delim("data/sdg_08/new/sdg_08_11.tsv", delim = ",")
sdg_08.sdg_08_20 <- read_delim("data/sdg_08/new/sdg_08_20.tsv", delim = ",")
sdg_08.sdg_08_20a <- read_delim("data/sdg_08/new/sdg_08_20a.tsv", delim = ",")
sdg_08.sdg_08_30 <- read_delim("data/sdg_08/new/sdg_08_30.tsv", delim = ",")
sdg_08.sdg_08_30a <- read_delim("data/sdg_08/new/sdg_08_30a.tsv", delim = ",")
sdg_08.sdg_08_40 <- read_delim("data/sdg_08/new/sdg_08_40.tsv", delim = ",")
sdg_08.sdg_08_60 <- read_delim("data/sdg_08/new/sdg_08_60.tsv", delim = ",")
sdg_09.sdg_04_20 <- read_delim("data/sdg_09/new/sdg_04_20.tsv", delim = ",")
sdg_09.sdg_09_10 <- read_delim("data/sdg_09/new/sdg_09_10.tsv", delim = ",")
sdg_09.sdg_09_30 <- read_delim("data/sdg_09/new/sdg_09_30.tsv", delim = ",")
sdg_09.sdg_09_40 <- read_delim("data/sdg_09/new/sdg_09_40.tsv", delim = ",")
sdg_09.sdg_09_50 <- read_delim("data/sdg_09/new/sdg_09_50.tsv", delim = ",")
sdg_09.sdg_09_60 <- read_delim("data/sdg_09/new/sdg_09_60.tsv", delim = ",")
sdg_09.sdg_09_70 <- read_delim("data/sdg_09/new/sdg_09_70.tsv", delim = ",")
sdg_09.sdg_12_61 <- read_delim("data/sdg_09/new/sdg_12_61.tsv", delim = ",")
sdg_09.sdg_17_60 <- read_delim("data/sdg_09/new/sdg_17_60.tsv", delim = ",")
sdg_10.sdg_01_10a <- read_delim("data/sdg_10/new/sdg_01_10a.tsv", delim = ",")
sdg_10.sdg_01_20a <- read_delim("data/sdg_10/new/sdg_01_20a.tsv", delim = ",")
sdg_10.sdg_04_10a <- read_delim("data/sdg_10/new/sdg_04_10a.tsv", delim = ",")
sdg_10.sdg_08_20a <- read_delim("data/sdg_10/new/sdg_08_20a.tsv", delim = ",")
sdg_10.sdg_08_30a <- read_delim("data/sdg_10/new/sdg_08_30a.tsv", delim = ",")
sdg_10.sdg_10_10 <- read_delim("data/sdg_10/new/sdg_10_10.tsv", delim = ",")
sdg_10.sdg_10_20 <- read_delim("data/sdg_10/new/sdg_10_20.tsv", delim = ",")
sdg_10.sdg_10_30 <- read_delim("data/sdg_10/new/sdg_10_30.tsv", delim = ",")
sdg_10.sdg_10_41 <- read_delim("data/sdg_10/new/sdg_10_41.tsv", delim = ",")
sdg_10.sdg_10_50 <- read_delim("data/sdg_10/new/sdg_10_50.tsv", delim = ",")
sdg_10.sdg_10_60 <- read_delim("data/sdg_10/new/sdg_10_60.tsv", delim = ",")
sdg_11.sdg_06_20 <- read_delim("data/sdg_11/new/sdg_06_20.tsv", delim = ",")
sdg_11.sdg_09_50 <- read_delim("data/sdg_11/new/sdg_09_50.tsv", delim = ",")
sdg_11.sdg_11_11 <- read_delim("data/sdg_11/new/sdg_11_11.tsv", delim = ",")
sdg_11.sdg_11_20 <- read_delim("data/sdg_11/new/sdg_11_20.tsv", delim = ",")
sdg_11.sdg_11_31 <- read_delim("data/sdg_11/new/sdg_11_31.tsv", delim = ",")
sdg_11.sdg_11_40 <- read_delim("data/sdg_11/new/sdg_11_40.tsv", delim = ",")
sdg_11.sdg_11_51 <- read_delim("data/sdg_11/new/sdg_11_51.tsv", delim = ",")
sdg_11.sdg_11_60 <- read_delim("data/sdg_11/new/sdg_11_60.tsv", delim = ",")
sdg_11.sdg_16_20 <- read_delim("data/sdg_11/new/sdg_16_20.tsv", delim = ",")
sdg_12.sdg_07_30 <- read_delim("data/sdg_12/new/sdg_07_30.tsv", delim = ",")
sdg_12.sdg_12_10 <- read_delim("data/sdg_12/new/sdg_12_10.tsv", delim = ",")
sdg_12.sdg_12_21 <- read_delim("data/sdg_12/new/sdg_12_21.tsv", delim = ",")
sdg_12.sdg_12_30 <- read_delim("data/sdg_12/new/sdg_12_30.tsv", delim = ",")
sdg_12.sdg_12_41 <- read_delim("data/sdg_12/new/sdg_12_41.tsv", delim = ",")
sdg_12.sdg_12_50 <- read_delim("data/sdg_12/new/sdg_12_50.tsv", delim = ",")
sdg_12.sdg_12_61 <- read_delim("data/sdg_12/new/sdg_12_61.tsv", delim = ",")
sdg_13.sdg_07_40 <- read_delim("data/sdg_13/new/sdg_07_40.tsv", delim = ",")
sdg_13.sdg_12_30 <- read_delim("data/sdg_13/new/sdg_12_30.tsv", delim = ",")
sdg_13.sdg_13_10 <- read_delim("data/sdg_13/new/sdg_13_10.tsv", delim = ",")
sdg_13.sdg_13_21 <- read_delim("data/sdg_13/new/sdg_13_21.tsv", delim = ",")
sdg_13.sdg_13_40 <- read_delim("data/sdg_13/new/sdg_13_40.tsv", delim = ",")
sdg_13.sdg_13_50 <- read_delim("data/sdg_13/new/sdg_13_50.tsv", delim = ",")
sdg_13.sdg_13_60 <- read_delim("data/sdg_13/new/sdg_13_60.tsv", delim = ",")
sdg_14.sdg_14_10 <- read_delim("data/sdg_14/new/sdg_14_10.tsv", delim = ",")
sdg_14.sdg_14_21 <- read_delim("data/sdg_14/new/sdg_14_21.tsv", delim = ",")
sdg_14.sdg_14_30 <- read_delim("data/sdg_14/new/sdg_14_30.tsv", delim = ",")
sdg_14.sdg_14_40 <- read_delim("data/sdg_14/new/sdg_14_40.tsv", delim = ",")
sdg_14.sdg_14_50 <- read_delim("data/sdg_14/new/sdg_14_50.tsv", delim = ",")
sdg_14.sdg_14_60 <- read_delim("data/sdg_14/new/sdg_14_60.tsv", delim = ",")
sdg_15.sdg_06_30 <- read_delim("data/sdg_15/new/sdg_06_30.tsv", delim = ",")
sdg_15.sdg_06_50 <- read_delim("data/sdg_15/new/sdg_06_50.tsv", delim = ",")
sdg_15.sdg_15_10 <- read_delim("data/sdg_15/new/sdg_15_10.tsv", delim = ",")
sdg_15.sdg_15_20 <- read_delim("data/sdg_15/new/sdg_15_20.tsv", delim = ",")
sdg_15.sdg_15_41 <- read_delim("data/sdg_15/new/sdg_15_41.tsv", delim = ",")
sdg_15.sdg_15_50 <- read_delim("data/sdg_15/new/sdg_15_50.tsv", delim = ",")
sdg_15.sdg_15_60 <- read_delim("data/sdg_15/new/sdg_15_60.tsv", delim = ",")
sdg_15.sdg_15_61 <- read_delim("data/sdg_15/new/sdg_15_61.tsv", delim = ",")
sdg_16.sdg_16_10 <- read_delim("data/sdg_16/new/sdg_16_10.tsv", delim = ",")
sdg_16.sdg_16_20 <- read_delim("data/sdg_16/new/sdg_16_20.tsv", delim = ",")
sdg_16.sdg_16_30 <- read_delim("data/sdg_16/new/sdg_16_30.tsv", delim = ",")
sdg_16.sdg_16_40 <- read_delim("data/sdg_16/new/sdg_16_40.tsv", delim = ",")
sdg_16.sdg_16_50 <- read_delim("data/sdg_16/new/sdg_16_50.tsv", delim = ",")
sdg_16.sdg_16_60 <- read_delim("data/sdg_16/new/sdg_16_60.tsv", delim = ",")
sdg_17.sdg_17_10 <- read_delim("data/sdg_17/new/sdg_17_10.tsv", delim = ",")
sdg_17.sdg_17_20 <- read_delim("data/sdg_17/new/sdg_17_20.tsv", delim = ",")
sdg_17.sdg_17_30 <- read_delim("data/sdg_17/new/sdg_17_30.tsv", delim = ",")
sdg_17.sdg_17_40 <- read_delim("data/sdg_17/new/sdg_17_40.tsv", delim = ",")
sdg_17.sdg_17_50 <- read_delim("data/sdg_17/new/sdg_17_50.tsv", delim = ",")
sdg_17.sdg_17_60 <- read_delim("data/sdg_17/new/sdg_17_60.tsv", delim = ",")


```

## Data cleaning and preprocessing

```{r}
# Get a list of all database objects in the environment
database_names <- ls(pattern = "sdg_(.*)")

# --------------Change name from geo/time to geo_time
# Function to change column name
change_column_name <- function(db_name) {
  db <- get(db_name)
  if ("geo\\time" %in% colnames(db)) {
    db <- db %>% rename(geo_time = `geo\\time`)
    assign(db_name, db, envir = .GlobalEnv)
    message(paste("Column name changed in", db_name))
  } else {
    message(paste("No column named 'geo\\time' found in", db_name))
  }
}

# Apply column name change to all databases
lapply(database_names, change_column_name)
```

```{r}
# Remove database without geo_time and select Spain and Germany
remove_db_without_geo_time <- function(db_name) {
  db <- get(db_name)
  if ("geo_time" %in% colnames(db)) {
    db <- db %>% 
      filter(geo_time %in% c("ES", "DE"))
    assign(db_name, db, envir = .GlobalEnv)
    message(paste("Filtered rows in", db_name))
  } else {
    message(paste("No column named 'geo_time' found in", db_name))
    rm(list = db_name, envir = .GlobalEnv)
    message(paste("Removed", db_name, "from the environment"))
  }
}

# Apply column filtering to all the databases
lapply(database_names, remove_db_without_geo_time)
```


```{r}
# Get a new list of all database objects in the environment
database_names <- ls(pattern = "sdg_(.*)")

# Chaage ":" to missing values (NAs)
change_colon_to_na <- function(db_name) {
  db <- get(db_name)
  
  if (is.data.frame(db) || is.matrix(db)) {
    db[] <- lapply(db, function(x) ifelse(grepl(":", x), NA, x))
    assign(db_name, db, envir = .GlobalEnv)
    message(paste("Changed ':' to NA in", db_name))
  } else if (is.list(db)) {
    db <- lapply(db, function(x) {
      if (is.data.frame(x) || is.matrix(x)) {
        x[] <- lapply(x, function(y) ifelse(y == ":", NA, y))
        message(paste("Changed ':' to NA in a nested data frame or matrix within", db_name))
      }
      x
    })
    assign(db_name, db, envir = .GlobalEnv)
  } else {
    message(paste("Unsupported data structure found in", db_name))
  }
}

# Apply column filtering to all the databases
lapply(database_names, change_colon_to_na)

```

There are some special characters that Eurostat uses to mark some data characteristics: 
-------------MISSING: Which ones---------------------

```{r}
#Remove this special characters
remove_chars_from_columns <- function(db_name) {
  db <- get(db_name)
  
  if (is.data.frame(db) || is.matrix(db)) {
    columns_to_process <- grep("^19|^20", colnames(db), value = TRUE)
    columns_to_process <- intersect(columns_to_process, colnames(db))
    
    db[, columns_to_process] <- lapply(db[, columns_to_process], function(x) gsub("[b-cdefnprsu-z]", "", x))
    
    assign(db_name, db, envir = .GlobalEnv)
    message(paste("Removed specified characters from columns starting with '19' or '20' in", db_name))
  } else if (is.list(db)) {
    db <- lapply(db, function(x) {
      if (is.data.frame(x) || is.matrix(x)) {
        columns_to_process <- grep("^19|^20", colnames(x), value = TRUE)
        columns_to_process <- intersect(columns_to_process, colnames(x))
        
        x[, columns_to_process] <- lapply(x[, columns_to_process], function(y) gsub("[b-cdefnprsu-z]", "", y))
        
        message(paste("Removed specified characters from columns starting with '19' or '20' in a nested data frame or matrix within", db_name))
      }
      x
    })
    assign(db_name, db, envir = .GlobalEnv)
  } else {
    message(paste("Unsupported data structure found in", db_name))
  }
}

# Apply the function to all databases
lapply(database_names, remove_chars_from_columns)

```

**Combining indicators from Spain and Germany in 2 datasets**

Reasons: 

- Increased data size: Combining indicators from multiple goals and countries will result in a larger dataset, providing more data points for the clustering algorithm. Having more data can improve the quality and robustness of the clustering results.

- Enhanced pattern discovery: Clustering algorithms often perform better with a larger number of data points. By combining indicators, you may uncover additional patterns and relationships that might be missed when clustering smaller subsets of indicators.

In order to perform the analysis and to create the final datasets we will need to remove all the non essential columns, as units and other specifications. 

```{r}
remove_columns <- function(database_names) {
  for (database_name in database_names) {
    database <- get(database_name)

    # Identify columns to remove
    columns_to_remove <- c()
    for (col_name in colnames(database)) {
      if (!(col_name == "geo_time" || col_name == "indicator" || grepl("^20", col_name))) { #we also remove years before 2000 since there is no information -------------> HAY QUE COMPROBAR ESTO
        columns_to_remove <- c(columns_to_remove, col_name)
      }
    }

    # Remove columns
    database <- database[, !(colnames(database) %in% columns_to_remove)]

    # Add updated database to global environment
    assign(database_name, database, envir = .GlobalEnv)
  }
}
#Apply the function to all the databases
remove_columns(database_names)
```



```{r}
# Create dataset for Spain
spain_dataset <- data.frame()
spain_database_names <- ls(pattern = "^sdg_")
for (database_name in spain_database_names) {
  database <- get(database_name)
  database$geo_time <- as.character(database$geo_time)  # Convert geo_time column to character
  database$indicator <- as.character(database$indicator) 
  spain_rows <- database[database$geo_time == "ES", ]
  spain_dataset <- bind_rows(spain_dataset, spain_rows)
}

# Create dataset for Germany
germany_dataset <- data.frame()
germany_database_names <- ls(pattern = "^sdg_")
for (database_name in germany_database_names) {
  database <- get(database_name)
  database$geo_time <- as.character(database$geo_time)  # Convert geo_time column to character
  database$indicator <- as.character(database$indicator)
  germany_rows <- database[database$geo_time == "DE", ]
  germany_dataset <- bind_rows(germany_dataset, germany_rows)
}

```


```{r}
# Convert to numeric
spain_dataset[, -1:-2] <- lapply(spain_dataset[, -1:-2], function(x) as.numeric(as.character(x)))
germany_dataset[, -1:-2] <- lapply(germany_dataset[, -1:-2], function(x) as.numeric(as.character(x)))
```


```{r}
# Remove the column for 2022 in Spain dataset
spain_dataset <- spain_dataset %>% select(-`2022 `)
# Remove the column for 2022 in Germany dataset
germany_dataset <- germany_dataset %>% select(-`2022 `)

```


```{r}
# Reorder columns in Spain dataset in reverse chronological order
spain_dataset <- spain_dataset[, rev(order(colnames(spain_dataset)))]
# Reorder columns in Germany dataset in reverse chronological order
germany_dataset <- germany_dataset[, rev(order(colnames(germany_dataset)))]
```


The dataset will be changed into long format in order to apply better analysis techniques

```{r}
#creating the new datasets
spain_trends <- spain_dataset %>%
  pivot_longer(cols = starts_with("20"), 
               names_to = "year", 
               values_to = "value")

germany_trends <- germany_dataset %>%
  pivot_longer(cols = starts_with("20"), 
               names_to = "year", 
               values_to = "value")
```



```{r}
#save the datsets 
write.csv(spain_trends, file = "./data/spain_trends.csv")
write.csv(germany_trends, file = "./data/germany_trends.csv")
```

