---
title: "Draft TFM"
author: "Victoria Costas"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
---

```{r global_options, include=T, echo = F}
knitr::opts_chunk$set(echo = T, warning=FALSE, message=FALSE)
```

```{=html}
<style>
body {
text-align: justify}
</style>
```

######## CLUSTERING PROBLEMS

- Clustering cannot be done with NAs, so how to proceed?
    - Imputing with mice (doesn't run and not correct due to non random missingness)
    - Imputing with mean (not the same distribution)
    - Remove rows (very few observations)
- Cannot due clustering of one year since I need more than one column, normally more than 10. 


## Libraries

```{r}
library(tidyverse)
library(dplyr)
library(mice)
library(factoextra)
library(zoo)
```

## Read the data

```{r}
spain <- read_csv("./data/spain_dataset.csv")
germany <- read_csv("./data/germany_dataset.csv")

#remove the first row for clarity
spain <- spain[, -1]
germany <- germany[, -1]
```

## Clustering analysis

```{r}

```


### Descriptive analysis

Some preliminar analysis:

```{r}
summary(spain)
summary(germany)
```

## Some pre-processing

```{r}
#Check NAs: 
# Calculate percentage of missing values in each column
missing_percentage_germany <- sapply(germany, function(x) {
  sum(is.na(x)) / length(x) * 100
})
# Create a data frame with column names and missing percentages
missing_data_germany <- data.frame(column = names(germany), missing_percentage = missing_percentage_germany)
# Sort the data frame by missing percentages in descending order
missing_data_germany <- missing_data_germany[order(missing_data_germany$missing_percentage, decreasing = TRUE), ]

# Calculate percentage of missing values in each column
missing_percentage_spain <- sapply(spain, function(x) {
  sum(is.na(x)) / length(x) * 100
})
# Create a data frame with column names and missing percentages
missing_data_spain <- data.frame(column = names(spain), missing_percentage = missing_percentage_spain)
# Sort the data frame by missing percentages in descending order
missing_data_spain <- missing_data_spain[order(missing_data_spain$missing_percentage, decreasing = TRUE), ]

missing_data_germany
missing_data_spain
```

We will do the first analysis for only one particular period in this case 2016, since there are quality data in this period and was when SDG started. -----------FALTA EXPLICACIÓN 

### Clustering

```{r}
#remove columns with more than 60% of missing values
ES <- spain %>% select(-c("2001", "2002","2003"))
DE <- germany %>% select(-c("2000","2001", "2002","2003"))  
  
ES$`2021` <- na.aggregate(ES$`2021`, FUN = mean)  
ES$`2020` <- na.aggregate(ES$`2020`, FUN = mean)
ES$`2019` <- na.aggregate(ES$`2019`, FUN = mean)
ES$`2018` <- na.aggregate(ES$`2018`, FUN = mean)
ES$`2017` <- na.aggregate(ES$`2017`, FUN = mean)
ES$`2016` <- na.aggregate(ES$`2016`, FUN = mean)
ES$`2015` <- na.aggregate(ES$`2015`, FUN = mean)
ES$`2000` <- na.aggregate(ES$`2000`, FUN = mean)
ES$`2004` <- na.aggregate(ES$`2004`, FUN = mean)
ES$`2005` <- na.aggregate(ES$`2005`, FUN = mean)
ES$`2006` <- na.aggregate(ES$`2006`, FUN = mean)
ES$`2007` <- na.aggregate(ES$`2007`, FUN = mean)
ES$`2008` <- na.aggregate(ES$`2008`, FUN = mean)
ES$`2009` <- na.aggregate(ES$`2009`, FUN = mean)
ES$`2010` <- na.aggregate(ES$`2010`, FUN = mean)
ES$`2011` <- na.aggregate(ES$`2011`, FUN = mean)
ES$`2012` <- na.aggregate(ES$`2012`, FUN = mean)
ES$`2013` <- na.aggregate(ES$`2013`, FUN = mean)
ES$`2014` <- na.aggregate(ES$`2014`, FUN = mean)

DE$`2021` <- na.aggregate(DE$`2021`, FUN = mean)
DE$`2020` <- na.aggregate(DE$`2020`, FUN = mean)
DE$`2019` <- na.aggregate(DE$`2019`, FUN = mean)
DE$`2018` <- na.aggregate(DE$`2018`, FUN = mean)
DE$`2017` <- na.aggregate(DE$`2017`, FUN = mean)
DE$`2016` <- na.aggregate(DE$`2016`, FUN = mean)
DE$`2015` <- na.aggregate(DE$`2015`, FUN = mean)
DE$`2004` <- na.aggregate(DE$`2004`, FUN = mean)
DE$`2005` <- na.aggregate(DE$`2005`, FUN = mean)
DE$`2006` <- na.aggregate(DE$`2006`, FUN = mean)
DE$`2007` <- na.aggregate(DE$`2007`, FUN = mean)
DE$`2008` <- na.aggregate(DE$`2008`, FUN = mean)
DE$`2009` <- na.aggregate(DE$`2009`, FUN = mean)
DE$`2010` <- na.aggregate(DE$`2010`, FUN = mean)
DE$`2011` <- na.aggregate(DE$`2011`, FUN = mean)
DE$`2012` <- na.aggregate(DE$`2012`, FUN = mean)
DE$`2013` <- na.aggregate(DE$`2013`, FUN = mean)
DE$`2014` <- na.aggregate(DE$`2014`, FUN = mean)
```

```{r}
#################NO VA###############
merged <- bind_rows(spain, germany)
#First we need to take only 2016 in Spain and Germany
data_2016 <-  merged %>% select(c("2016", geo_time))

data_impute <- merged[, c("geo_time", "2015", "2016", "2017", "2018", "2019", "2020")]
data_impute$geo_time <- as.factor(data_impute$geo_time)
# Create the formula for imputation
formula <- as.formula(paste("Cbind(2015, 2016, 2017, 2018, 2019, 2020)", "~ geo_time"))
# Perform imputation
mice_mod <- mice(data_impute, method = c("mean"), m = 1, formula = list(formula))
imputed_data <- complete(mice_mod)
```

```{r}
m = 4 
mice_mod <- mice(data_2016) 
df <- complete(mice_mod, action=m)
```


Initial guess for the number of clusters: 7  
  
```{r}
ES <- ES %>% select(-c(geo_time))
DE <- DE %>% select(-c(geo_time))
set.seed(123)
k = 3
fit <- kmeans(scale(ES), centers = k, nstart = 10000)
fit <- kmeans(scale(DE), centers = k, nstart = 10000)

groups <- fit$cluster
barplot(table(groups), col = "#FF00FF")

fviz_nbclust(scale(ES), kmeans, method = 'wss', k.max = 10, nstart = 1000) 
#optimal number of clusters 3 

cluster_labels <- fit$cluster
```
## Interpretation of centers

```{r}
centers=fit$centers

barplot(centers[1,], las=2, col="#FF00FF")
barplot(centers[2,], las=2, col="#FF00FF")
barplot(centers[3,], las=2, col="#FF00FF")

```

## Clusplot 

```{r}
fviz_cluster(fit, data = ES, geom = c("point"),ellipse.type = 'norm', pointsize=1)+
  theme_minimal()+scale_fill_brewer(palette="Paired")

fviz_cluster(fit, data = DE, geom = c("point"),ellipse.type = 'norm', pointsize=1)+
  theme_minimal()+scale_fill_brewer(palette="Paired")
```

## Number of groups

We will use different methods (cluster sums of squares, average silhouette and gap statistics) to know which is the optimal number of clusters:

```{r}
fviz_nbclust(scale(X), kmeans, method = 'wss', k.max = 10, nstart = 1000) 
fviz_nbclust(scale(X), kmeans, method = 'silhouette', k.max = 10, nstart = 1000)  
fviz_nbclust(scale(X), kmeans, method = 'gap_stat', k.max = 10, nstart = 100, nboot = 500) 
```

Despite the fact that the number of ideal clusters indicated to us is not seven, in this analysis we will leave this figure because the number of regions is quite high and in this way we do not saturate the clusters.

```{r}
fit.km <-eclust(data_2016, "kmeans", stand = T, k = 3, graph = T)
```


```{r}

```


```{r}

```


```{r}

```


```{r}

```





















