---
title: "Factor Analysis"
author: "Victoria Costas"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
---

```{r global_options, include=T, echo = F}
knitr::opts_chunk$set(echo = T, warning=FALSE, message=FALSE)
```

```{=html}
<style>
body {
text-align: justify}
</style>
```

## Libraries

```{r}
library(tidyverse)
library(dplyr)
library(mice)
library(missForest)
library(ggplot2)
library(polycor)
library(ggcorrplot)
library(stats)
library(psych)
library(GPArotation)
```

## Read the data

```{r}
data <- read_csv("./data/data_mf.csv")

#remove the first column for clarity
data <- data[, -1]
```

### Descriptive analysis

Some preliminary analysis:

```{r}
summary(data)
```

## Factor analysis

The steps for conducting a factor analysis are summarised as follows:
1. Verification that the matrix is factorisable
2. Extraction of factors
3. Factor rotation
4. Interpretation of results 

```{r}
data <- data[, -1:-2]
corr <- cor(data)
ggcorrplot(corr, hc.order = T, type="lower")
```

The above graph is the correlation matrix that shows the relationship between the different goals. Once we have verified that there is a certain correlation between them, we will use Bartlett's test of sphericity and the Kaiser-Meyer-Olkin test to check that our matrix is indeed factorisable.

```{r}
test <- cortest.bartlett(corr)
test$p
KMO(corr)
```

First we will perform Bartlett's test which would test the null hypothesis that the variables are uncorrelated in the population. In this case the p-value is much lower than 0.05 so we can say that we reject the null hypothesis and there is enough correlation to use factor analysis. Secondly, we will apply the Kaiser-Meyer-Olkin criterion which also indicates whether the data are suitable for factor analysis. In this case the overall MSA is 0.74 which means that the data are suitable for factor analysis.

```{r}
#models with 4 factors
model_mle<-fa(corr,
           nfactors = 4,
           rotate = "none",
           fm="mle") #maximum likelihood model

model_minres<-fa(corr,
           nfactors = 4,
           rotate = "none",
           fm="minres") #minimum residual model
```


```{r}
#comparing communalities
c1 <- sort(model_mle$communality, decreasing = T) 
c2 <- sort(model_minres$communality, decreasing = T)
head(cbind(c1,c2))
```

Communalities represent the proportion of variance in each observed variable that is accounted for by the extracted factors. The values indicate how well the factors explain the corresponding SDGs. Higher values suggest that the factors capture a larger portion of the variance in the SDGs.

```{r}
#comparison of uniquenesses  
u1 <- sort(model_mle$uniquenesses,decreasing = T)
u2 <- sort(model_minres$uniquenesses,decreasing = T)
head(cbind(u1,u2))
```

Uniquenesses, on the other hand, represent the proportion of variance in each observed variable that is unique or not explained by the factors. Higher uniqueness values indicate that the corresponding SDGs are less explained by the extracted factors. If we look at the results of both models, they seem to perform quite similarly in both communalities and uniquenesses. These results indicate that both factor models capture and explain a considerable portion of the variance in the SDGs.

```{r}
scree(corr)
```

```{r}
fa.parallel(corr,n.obs=200,fa="fa",fm="minres")
```

The above graphs serve as a guide to identify the number of factors to be used in the models, in this case the parallel analysis indicates the ideal number of factors is 5. 

```{r}
#Rotation
rot<-c("none", "varimax", "quartimax","Promax")

bi_mod<-function(type){
biplot.psych(fa(data, nfactors = 2, fm ="minres", rotate = type), main = paste("Biplot with rotation", type), col=c(2,3,4), pch = c(21,18))  
}
sapply(rot,bi_mod)
```

Rotation serves as a technique to simplify and interpret the factor structure obtained from the initial factor extraction. The rotation aims to achieve a more interpretable and meaningful solution by maximizing the clarity of the factor loadings. For the interpretation, we use varimax rotation in order to facilitate the interpretability of the factors since this method simplifies the matrix and interpreting factor analysis is generally difficult.

```{r}
model_varimax <- fa(corr, nfactors = 5, rotate = "varimax", fa="minres")

fa.diagram(model_varimax)
```

Each factor represents a group of variables (SDGs) that share common variance:

- MR1: Includes Goal 17, Goal 16, Goal 6, Goal 9, Goal 15, and Goal 14. These SDGs are strongly associated with MR1, suggesting that they share common underlying factors and are interrelated. The strength of the relationship between these SDGs and MR1 is likely high.

- MR2: Includes Goal 13, Goal 2, and Goal 1. Among these SDGs, Goal 13 has a unique relationship with MR2, suggesting that it has a distinct association with this factor. Goals 2 and 1 are also associated with MR2, although the nature of their relationship may be different from that of Goal 13.

- MR3: Includes Goal 11, Goal 3, and Goal 12. These SDGs are associated with MR3, indicating that they share common factors. The strength of the relationship between these SDGs and MR3 is likely high.

- MR4: Includes Goal 10, Goal 8, and Goal 5. These SDGs are grouped together and associated with MR4, suggesting that they share underlying factors. The strength of the relationship between these SDGs and MR4 is likely high.

- MR5: Includes Goal 4. This SDG is uniquely associated with MR5, indicating that it has a distinct relationship with this factor.

Goal 7 is not assigned to any specific factor, suggesting that it does not show a strong relationship with any of the five factors identified in the factor analysis, irrespective of the number of factors used in the analysis.

```{r}
print(model_varimax$loadings, cut = 0) 
model_varimax$uniquenesses
model_varimax$communalities
```

In these last results we see a summary of the factor analysis, showing the relationships between the goals and the factors (loadings), also showing the strength of the factors by the sum of squares loadings. Finally, s variables have unique relationships with specific factors. For example, Goal 13 has a strong loading on MR2, suggesting a distinct relationship between Goal 13 and this factor as stated before. 

### Factor scores for cluster analysis

```{r}
# Obtain factor scores
factor_scores <- predict(model_varimax, data = data)

# Perform cluster analysis using factor scores
cluster_results <- kmeans(factor_scores, centers = 7)

# View cluster membership for each observation
cluster_membership <- cluster_results$cluster

# View cluster centroids
cluster_centroids <- cluster_results$centers

# Define the unique cluster labels
unique_clusters <- unique(cluster_membership)

# Create a color palette for the clusters
color_palette <- rainbow(length(unique_clusters))

# Plotting the cluster analysis results
plot(factor_scores, col = color_palette[cluster_membership], pch = 16, main = "Cluster Analysis Results")
# Add a legend
legend("topright", legend = unique_clusters, col = color_palette, pch = 16, title = "Clusters")
```














