---
title: "Simple Trends Analysis"
author: "Victoria Costas"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
---

```{r global_options, include=T, echo = F}
knitr::opts_chunk$set(echo = T, warning=FALSE, message=FALSE)
```

```{=html}
<style>
body {
text-align: justify}
</style>
```

# Trends analysis 

This simple trend analysis document is only a support for the final dashboard, this means that the graphs here will be included in the final dashboard but they are included in this document in order to have them collected, commented and analysed. 


## Libraries

```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(patchwork)
```

## Read the data

```{r}
spain_trends <- read_csv("./data/spain_trends.csv")
germany_trends <- read_csv("./data/germany_trends.csv")

#remove the first row for clarity
spain_trends <- spain_trends[, -1]
germany_trends <- germany_trends[, -1]
```


## Normalize the data

```{r}
trends <- rbind(spain_trends, germany_trends)

# Scale the "value" column in the "trends" dataset
trends_scaled <- trends 
trends_scaled$value <- scale(trends$value)

# View the scaled dataset
head(trends_scaled)
```


Once scaled the data, we will compute the average for each goal.

```{r}
#compute the average for each goal each year
average_by_goal_year <- trends_scaled %>%
  group_by(indicator, year, geo_time) %>%
  summarise(Average_Value = mean(value, na.rm = TRUE))

head(average_by_goal_year)
```


## Trends analysis

### Simple analysis

We will merge the data in order to perform the analysis

```{r}
#data for the year 2010 onwards
data <- average_by_goal_year %>% filter(year >= 2010)
```


```{r}
# Create line plots for each goal
plot1 <- data %>% filter(indicator == "ind1") %>% 
  ggplot(aes(x = year, y = Average_Value, color = geo_time, group = geo_time)) +
  geom_line(size = 1.5) +
  labs(title = "Progress of Goal 1 (No Poverty)", x = "Year", y = "Average Value") +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, face = "bold"),
        axis.title = element_text(size = 12),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
        axis.text.y = element_text(size = 10)) +
  scale_color_manual(values = c("ES" = "blue", "DE" = "red"),
                     labels = c("ES" = "Spain", "DE" = "Germany"),
                     name = "Country")

plot2 <- data %>% filter(indicator == "ind2") %>% 
  ggplot(aes(x = year, y = Average_Value, color = geo_time, group = geo_time)) +
  geom_line(size = 1.5) +
  labs(title = "Progress of Goal 2 (Zero Hunger)", x = "Year", y = "Average Value") +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, face = "bold"),
        axis.title = element_text(size = 12),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
        axis.text.y = element_text(size = 10),
        legend.position = "none")+
  scale_color_manual(values = c("ES" = "blue", "DE" = "red"),
                     labels = c("ES" = "Spain", "DE" = "Germany"),
                     name = "Country")

plot3 <- data %>% filter(indicator == "ind3") %>% 
  ggplot(aes(x = year, y = Average_Value, color = geo_time, group = geo_time)) +
  geom_line(size = 1.5) +
  labs(title = "Progress of Goal 3 (Good Health and Well-being)", x = "Year", y = "Average Value") +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, face = "bold"),
        axis.title = element_text(size = 12),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
        axis.text.y = element_text(size = 10),
        legend.position = "none")+
  scale_color_manual(values = c("ES" = "blue", "DE" = "red"),
                     labels = c("ES" = "Spain", "DE" = "Germany"),
                     name = "Country")

plot4 <- data %>% filter(indicator == "ind4") %>% 
  ggplot(aes(x = year, y = Average_Value, color = geo_time, group = geo_time)) +
  geom_line(size = 1.5) +
  labs(title = "Progress of Goal 4 (Quality Education)", x = "Year", y = "Average Value") +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, face = "bold"),
        axis.title = element_text(size = 12),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
        axis.text.y = element_text(size = 10),
        legend.position = "none")+
  scale_color_manual(values = c("ES" = "blue", "DE" = "red"),
                     labels = c("ES" = "Spain", "DE" = "Germany"),
                     name = "Country")

plot5 <- data %>% filter(indicator == "ind5") %>% 
  ggplot(aes(x = year, y = Average_Value, color = geo_time, group = geo_time)) +
  geom_line(size = 1.5) +
  labs(title = "Progress of Goal 5", x = "Year", y = "Average Value") +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, face = "bold"),
        axis.title = element_text(size = 12),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
        axis.text.y = element_text(size = 10),
        legend.position = "none")+
  scale_color_manual(values = c("ES" = "blue", "DE" = "red"),
                     labels = c("ES" = "Spain", "DE" = "Germany"),
                     name = "Country")

plot6 <- data %>% filter(indicator == "ind6") %>% 
  ggplot(aes(x = year, y = Average_Value, color = geo_time, group = geo_time)) +
  geom_line(size = 1.5) +
  labs(title = "Progress of Goal 6", x = "Year", y = "Average Value") +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, face = "bold"),
        axis.title = element_text(size = 12),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
        axis.text.y = element_text(size = 10),
        legend.position = "none")+
  scale_color_manual(values = c("ES" = "blue", "DE" = "red"),
                     labels = c("ES" = "Spain", "DE" = "Germany"),
                     name = "Country")

plot7 <- data %>% filter(indicator == "ind7") %>% 
  ggplot(aes(x = year, y = Average_Value, color = geo_time, group = geo_time)) +
  geom_line(size = 1.5) +
  labs(title = "Progress of Goal 7", x = "Year", y = "Average Value") +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, face = "bold"),
        axis.title = element_text(size = 12),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
        axis.text.y = element_text(size = 10),
        legend.position = "none")+
  scale_color_manual(values = c("ES" = "blue", "DE" = "red"),
                     labels = c("ES" = "Spain", "DE" = "Germany"),
                     name = "Country")

plot8 <- data %>% filter(indicator == "ind8") %>% 
  ggplot(aes(x = year, y = Average_Value, color = geo_time, group = geo_time)) +
  geom_line(size = 1.5) +
  labs(title = "Progress of Goal 8", x = "Year", y = "Average Value") +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, face = "bold"),
        axis.title = element_text(size = 12),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
        axis.text.y = element_text(size = 10),
        legend.position = "none")+
  scale_color_manual(values = c("ES" = "blue", "DE" = "red"),
                     labels = c("ES" = "Spain", "DE" = "Germany"),
                     name = "Country")

plot9 <- data %>% filter(indicator == "ind9") %>% 
  ggplot(aes(x = year, y = Average_Value, color = geo_time, group = geo_time)) +
  geom_line(size = 1.5) +
  labs(title = "Progress of Goal 9", x = "Year", y = "Average Value") +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, face = "bold"),
        axis.title = element_text(size = 12),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
        axis.text.y = element_text(size = 10),
        legend.position = "none")+
  scale_color_manual(values = c("ES" = "blue", "DE" = "red"),
                     labels = c("ES" = "Spain", "DE" = "Germany"),
                     name = "Country")

plot10 <- data %>% filter(indicator == "ind10") %>% 
  ggplot(aes(x = year, y = Average_Value, color = geo_time, group = geo_time)) +
  geom_line(size = 1.5) +
  labs(title = "Progress of Goal 10", x = "Year", y = "Average Value") +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, face = "bold"),
        axis.title = element_text(size = 12),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
        axis.text.y = element_text(size = 10),
        legend.position = "none")+
  scale_color_manual(values = c("ES" = "blue", "DE" = "red"),
                     labels = c("ES" = "Spain", "DE" = "Germany"),
                     name = "Country")

plot11 <- data %>% filter(indicator == "ind11") %>% 
  ggplot(aes(x = year, y = Average_Value, color = geo_time, group = geo_time)) +
  geom_line(size = 1.5) +
  labs(title = "Progress of Goal 11", x = "Year", y = "Average Value") +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, face = "bold"),
        axis.title = element_text(size = 12),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
        axis.text.y = element_text(size = 10),
        legend.position = "none")+
  scale_color_manual(values = c("ES" = "blue", "DE" = "red"),
                     labels = c("ES" = "Spain", "DE" = "Germany"),
                     name = "Country")

plot12 <- data %>% filter(indicator == "ind12") %>% 
  ggplot(aes(x = year, y = Average_Value, color = geo_time, group = geo_time)) +
  geom_line(size = 1.5) +
  labs(title = "Progress of Goal 12", x = "Year", y = "Average Value") +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, face = "bold"),
        axis.title = element_text(size = 12),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
        axis.text.y = element_text(size = 10),
        legend.position = "none")+
  scale_color_manual(values = c("ES" = "blue", "DE" = "red"),
                     labels = c("ES" = "Spain", "DE" = "Germany"),
                     name = "Country")

plot13 <- data %>% filter(indicator == "ind13") %>% 
  ggplot(aes(x = year, y = Average_Value, color = geo_time, group = geo_time)) +
  geom_line(size = 1.5) +
  labs(title = "Progress of Goal 13", x = "Year", y = "Average Value") +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, face = "bold"),
        axis.title = element_text(size = 12),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
        axis.text.y = element_text(size = 10),
        legend.position = "none")+
  scale_color_manual(values = c("ES" = "blue", "DE" = "red"),
                     labels = c("ES" = "Spain", "DE" = "Germany"),
                     name = "Country")

plot14 <- data %>% filter(indicator == "ind14") %>% 
  ggplot(aes(x = year, y = Average_Value, color = geo_time, group = geo_time)) +
  geom_line(size = 1.5) +
  labs(title = "Progress of Goal 14", x = "Year", y = "Average Value") +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, face = "bold"),
        axis.title = element_text(size = 12),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
        axis.text.y = element_text(size = 10),
        legend.position = "none")+
  scale_color_manual(values = c("ES" = "blue", "DE" = "red"),
                     labels = c("ES" = "Spain", "DE" = "Germany"),
                     name = "Country")

plot15 <- data %>% filter(indicator == "ind15") %>% 
  ggplot(aes(x = year, y = Average_Value, color = geo_time, group = geo_time)) +
  geom_line(size = 1.5) +
  labs(title = "Progress of Goal 15", x = "Year", y = "Average Value") +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, face = "bold"),
        axis.title = element_text(size = 12),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
        axis.text.y = element_text(size = 10),
        legend.position = "none")+
  scale_color_manual(values = c("ES" = "blue", "DE" = "red"),
                     labels = c("ES" = "Spain", "DE" = "Germany"),
                     name = "Country")

plot16 <- data %>% filter(indicator == "ind16") %>% 
  ggplot(aes(x = year, y = Average_Value, color = geo_time, group = geo_time)) +
  geom_line(size = 1.5) +
  labs(title = "Progress of Goal 16", x = "Year", y = "Average Value") +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, face = "bold"),
        axis.title = element_text(size = 12),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
        axis.text.y = element_text(size = 10),
        legend.position = "none")+
  scale_color_manual(values = c("ES" = "blue", "DE" = "red"),
                     labels = c("ES" = "Spain", "DE" = "Germany"),
                     name = "Country")

plot17 <- data %>% filter(indicator == "ind17") %>% 
  ggplot(aes(x = year, y = Average_Value, color = geo_time, group = geo_time)) +
  geom_line(size = 1.5) +
  labs(title = "Progress of Goal 17", x = "Year", y = "Average Value") +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, face = "bold"),
        axis.title = element_text(size = 12),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
        axis.text.y = element_text(size = 10),
        legend.position = "none")+
  scale_color_manual(values = c("ES" = "blue", "DE" = "red"),
                     labels = c("ES" = "Spain", "DE" = "Germany"),
                     name = "Country")

plots1_4 <- plot1 + plot2 + plot3 + plot4
plots5_8 <- plot5 + plot6 + plot7 + plot8
plots9_12 <- plot9 + plot10 + plot11 + plot12
plots13_16 <- plot13 + plot14 + plot15 + plot16

plots1_4
plots5_8
plots9_12
plots13_16
plot17
```


```{r}
summary_stats_2010 <- data %>%
  group_by(indicator, geo_time) %>%
  summarise(Average_Value = mean(Average_Value, na.rm = TRUE, Median_Value = median(Average_Value, na.rm = TRUE)))

print(summary_stats_2010)

plot <- summary_stats_2010 %>%
  ggplot(aes(x = reorder(indicator, -Average_Value), y = Average_Value, fill = geo_time)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Summary Statistics from 2010 to 2021",
       x = "Indicator",
       y = "Average Value",
       fill = "Country") +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, face = "bold"),
        axis.title = element_text(size = 12),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 10))

# Display the plot
plot
```


```{r}
growth <- data %>%
  arrange(indicator, geo_time, year) %>%
  group_by(indicator, geo_time) %>%
  filter(complete.cases(Average_Value)) %>%
  mutate(Growth_Rate = (Average_Value - lag(Average_Value)) / lag(Average_Value) * 100)

print(growth)

growth_plot_es <- growth %>% filter(geo_time == "ES") %>% 
  ggplot(aes(x = year, y = Growth_Rate, color = indicator, group = interaction(indicator, geo_time))) +
  geom_line() +
  geom_smooth(se = FALSE) +
  labs(title = "Growth Rate by Indicator and Year",
       x = "Year",
       y = "Growth Rate",
       color = "Indicator") +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, face = "bold"),
        axis.title = element_text(size = 12),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
        axis.text.y = element_text(size = 10))

growth_plot_es

growth_plot_de <- growth %>% filter(geo_time == "DE") %>% 
  ggplot(aes(x = year, y = Growth_Rate, color = indicator, group = interaction(indicator, geo_time))) +
  geom_line() +
  geom_smooth(se = FALSE) +
  labs(title = "Growth Rate by Indicator",
       x = "Year",
       y = "Growth Rate",
       color = "Indicator") +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, face = "bold"),
        axis.title = element_text(size = 12),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
        axis.text.y = element_text(size = 10))

growth_plot_de
```



```{r}
overall_stats_es <- data %>% filter(geo_time== "ES") %>% 
  group_by(indicator, year) %>%
  summarise(Average_Value = Average_Value, Max_Value = max(Average_Value, na.rm = T), Min_Value = min(Average_Value, na.rm = T))

print(overall_stats_es)

overall_stats_de <- data %>% filter(geo_time == "DE") %>% 
  group_by(indicator, year) %>%
  summarise(Average_Value = Average_Value, Max_Value = max(Average_Value, na.rm = T), Min_Value = min(Average_Value, na.rm = T))

print(overall_stats_de)
```


```{r}
# Combine overall_stats_es and overall_stats_de into a single data frame
overall_stats_combined <- rbind(overall_stats_es, overall_stats_de)

# Create bar plot
bar_plot <- ggplot(overall_stats_combined, aes(x = year, y = Average_Value, fill = indicator)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Overall Statistics: Spain vs Germany", x = "Year", y = "Average Value") +
  theme_minimal() +
  theme(legend.position = "top")

# Display the plot
bar_plot


###########FOR SPAIN
# Get unique years
unique_years <- unique(overall_stats_es$year)

# Create bar plots for each year
bar_plots <- lapply(unique_years, function(year) {
  data_year <- overall_stats_es[overall_stats_es$year == year, ]
  
  # Determine fill colors based on positive/negative values
  data_year$fill_color <- ifelse(data_year$Average_Value >= 0, "Positive", "Negative")
  
  ggplot(data_year, aes(x = indicator, y = Average_Value, fill = fill_color)) +
    geom_bar(stat = "identity", position = "dodge") +
    labs(title = paste("Overall Statistics:", year), x = "Indicator", y = "Average Value") +
    scale_fill_manual(values = c("Positive" = "lightgreen", "Negative" = "#FF6666")) +  # Set fill colors
    theme_minimal() +
    theme(legend.position = "none") +
    coord_flip()
})

# Display the bar plots one by one
for (i in seq_along(bar_plots)) {
  print(bar_plots[[i]])
}

###########FOR GERMANY
# Get unique years
unique_years <- unique(overall_stats_de$year)

# Create bar plots for each year
bar_plots <- lapply(unique_years, function(year) {
  data_year <- overall_stats_de[overall_stats_de$year == year, ]
  
  # Determine fill colors based on positive/negative values
  data_year$fill_color <- ifelse(data_year$Average_Value >= 0, "Positive", "Negative")
  
  ggplot(data_year, aes(x = indicator, y = Average_Value, fill = fill_color)) +
    geom_bar(stat = "identity", position = "dodge") +
    labs(title = paste("Overall Statistics:", year), x = "Indicator", y = "Average Value") +
    scale_fill_manual(values = c("Positive" = "lightgreen", "Negative" = "#FF6666")) +  # Set fill colors
    theme_minimal() +
    theme(legend.position = "none") +
    coord_flip()
})

# Display the bar plots one by one
for (i in seq_along(bar_plots)) {
  print(bar_plots[[i]])
}

########## FOR BOTH
# Get unique years
unique_years <- unique(overall_stats_combined$year)

# Create bar plots for each year
bar_plots <- lapply(unique_years, function(year) {
  data_year <- overall_stats_combined[overall_stats_combined$year == year, ]
  
  # Determine fill colors based on positive/negative values
  data_year$fill_color <- ifelse(data_year$Average_Value >= 0, "Positive", "Negative")
  
  ggplot(data_year, aes(x = indicator, y = Average_Value, fill = fill_color)) +
    geom_bar(stat = "identity", position = "dodge") +
    labs(title = paste("Overall Statistics:", year), x = "Indicator", y = "Average Value") +
    scale_fill_manual(values = c("Positive" = "lightgreen", "Negative" = "#FF6666")) +  # Set fill colors
    theme_minimal() +
    theme(legend.position = "none") +
    coord_flip()
})

# Display the bar plots one by one
for (i in seq_along(bar_plots)) {
  print(bar_plots[[i]])
}

```


